### 2013-11-04 presentation/ppt.tex
Presentation latex file.

\documentclass{beamer}
\mode<presentation> {
\usetheme{Boadilla}
\setbeamercovered{invisible}
\setbeamertemplate{navigation symbols}{} 
}

%\newcommand{\abs}[1]{\ensuremath{\left|#1\right|}}
%\newcommand{\bracket}[2]{\ensuremath{\left\langle#1 \vphantom{#2}\middle|  #2 \vphantom{#1}\right\rangle}}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\title[HCA]{Hierarchical Controller Architecture for SDN}
\author[]{\textbf{Gourav Khaneja\\Sweta Yamini Seethamraju} \\ \small{Supervisor:} Prof. Philip Brighten Godfrey}
\date{\today}

\begin{document}
\frame{
\titlepage
}

\frame{
\frametitle{Outline}
\tableofcontents
}

\section{Introduction}

\frame{
\frametitle{Problem}
\begin{itemize}
	\item Single controller architecture is not scalable
	\item Distributed controller architectures provide global view to control application
\pause
	\item Problems are
	\begin{itemize}
		\item Communication overhead
		\item Network state inconsistencies
		\item Management barriers across networks 
		\item Scalability of complete global view of the network
	\end{itemize}
\pause
	\item Trade-offs
	\begin{itemize}
		\item Consistency vs. Responsiveness
		\item Application design complexity (architecture aware vs. agnostic)
	\end{itemize}
\pause
	\item Aim
	\begin{itemize}
		\item Make control application agnostic to underlying distributed architecture
		\item Aggregate network to simplify network view
		\item Handle inconsistency in "SDN datapath"
	\end{itemize}
\end{itemize}
}

\frame{
\frametitle{Previous Work}
\begin{itemize}
	\item Hyperflow
\pause
	\begin{itemize}
		\item Flat distributed controller architecture
		\item Appliations are agnostic to underlying state distribution
		\item Provides a logically centralized global view of the network
		\item Results in performance degradation and transient inconsistencies
	\end{itemize}
\pause
	\item Onix
\pause
	\begin{itemize}
		\item Provides flexible framework to handle controller topology
		\item Provides generic distributed state management API
		\item Provides a framework but not an approach
		\item Design decisions left to control application
	\end{itemize}
\pause
	\item Kandoo
\pause
	\begin{itemize}
		\item Two-level controller architecture
		\item Events are propagated only on subscription
		\item Applications need to be aware of hierarchy
		\item View of root controller varies with control application
	\end{itemize}
\end{itemize}
}

\section{Approach}

\frame{
\frametitle{What }
\begin{itemize}
	\item
\end{itemize}
}

\frame{
\frametitle{Our Approach}
\begin{itemize}
	\item
\end{itemize}
}

\section{Progress}

\frame{
\frametitle{Progress}
\begin{itemize}
	\item
\end{itemize}
}

\frame{
\frametitle{Results}
\begin{itemize}
\item Preliminary link utilization plots
\end{itemize}
\begin{figure}
 \includegraphics[scale=0.2]{../chart_1}
\caption{Link utilization}
\end{figure}
}

\frame{
\frametitle{Future Work}
\begin{itemize}
	\item 
	\item Implement using OpenFlow
\end{itemize}
}

\frame{
\centerline{Thank you!}
}

\end{document}

### 2013-11-04 simulator/src/swiconsim/SwiConSim.java
flow rule installation is passed down the hierarchy

@@ -113,7 +113,7 @@ public class SwiConSim {
 		s2.addHost(h2, (short) 3);
 
 		// Add edge s1:3 <-> s2:2
-		DataNetwork.getInstance().addEdge(s1_id, (short) 2, s2_id, (short) 2);
+		DataNetwork.getInstance().addEdge(s1_id, (short) 3, s2_id, (short) 2);
 
 		System.out.println(c1.getTopology().toString());
 		System.out.println(c2.getTopology().toString());

### 2013-11-04 simulator/src/swiconsim/controller/Controller.java

@@ -55,6 +55,7 @@ public class Controller extends Node implements IControlPlane, IController,
 
 	@Override
 	public void addFlow(Flow flow) {
+		logger.info("**** " + flow.toString());
 		for (Long swid : nodes) {
 			addFlowToSwitch(swid, flow);
 		}
@@ -75,19 +76,6 @@ public class Controller extends Node implements IControlPlane, IController,
 	 */
 	@Override
 	public void sendPktInController(Packet pkt) {
-		// TODO Auto-generated method stub
-
-	}
-
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see
-	 * swiconsim.api.IControlPlane#receiveNotificationFromController(swiconsim
-	 * .messages.Message)
-	 */
-	@Override
-	public void receiveNotificationFromController(Message msg) {
 		// TODO Auto-generated method stub
 
 	}

### 2013-11-04 simulator/src/swiconsim/network/ManagementNetwork.java

@@ -102,6 +102,7 @@ public class ManagementNetwork implements IManagementNetwork, IDataNetwork {
 		logger.info(msg.getFrom() + "->" + msg.getTo() + " " + msg.getType());
 		long swid = msg.getTo();
 		if (nodeMap.containsKey(swid)) {
+			logger.info("Sending to " + swid);
 			nodeMap.get(swid).receiveNotificationFromController(msg);
 		} else {
 			logger.warning("Switch not found - " + swid);

### 2013-11-04 simulator/src/swiconsim/node/Node.java

@@ -78,6 +78,7 @@ public abstract class Node implements IControlPlane {
 
 	@Override
 	public void receiveNotificationFromController(Message msg) {
+		logger.info(" rcvd ");
 		switch (msg.getType()) {
 		case OFPFC_ADD:
 			Flow flow = (Flow) msg.getPayload();

### 2013-11-05 presentation/ppt.tex
Added result slide

@@
 \frame{
-\frametitle{Results}
-\begin{itemize}
-\item Preliminary link utilization plots
-\end{itemize}
+\frametitle{Preliminary Results}
+%\begin{itemize}
+%\item Preliminary link utilization plots
+%\end{itemize}
 \begin{figure}
- \includegraphics[scale=0.2]{../chart_1}
-\caption{Link utilization}
+ \includegraphics[scale=0.3]{chart_1}
+ \includegraphics[scale=0.3]{chart_2}
+\caption{Comparision of Link utilizations with/without Timestamp \footnote{ \scriptsize Simulator @ https://github.com/MugiwaraLuffy/ACN/tree/master/simulator/src/timestamp}}
 \end{figure}
+
+'Long lived flows' flow between domain 1 and domain 3 through domain 2. A link balancer control application runs on each controller
+
 }

### 2013-11-05 presentation/ppt.tex
approach

@@
 \begin{itemize}
-	\item
+	\item Hierarchical Architecture
+	\begin{itemize}
+		\item Controllers' view do not span the whole network, but only the 'underlying network'
+		\item Each controller behaves as a giant openflow switch for parent controller (Network Aggregation)
+		\item Controller translates OFP commands and network updates between parent controller and underlying networks
+	\end{itemize}
+	\item Network Aggregation
+	\begin{itemize}
+		\item Share a condensed view of the network: A Network flow table
+		\item Surpress network events such that aggregated network does not change: Reduce communication overhead
+		\item Convert flow addtion/removal commands on aggregated network to actual network in an efficient way
+	\end{itemize}
+	\item Timestamp: 
+		\item Handle domain level inconsistencies in "SDN Datapath" (Domain Network)
+		\item Each domain i maintains a vector timestamp T_i[1...n] 
+		\item (Routers in) A Domain may change flow depending on their timestamps
+	\end{itemize}   
 \end{itemize}
@@
 \frame{
-\frametitle{Progress}
+\frametitle{Network Aggregation}
+\begin{itemize}
+	\item
+\end{itemize}
+}
+
+\frame{
+\frametitle{Timestamp}
 \begin{itemize}
@@
 
+\section{What next ?}
+
 \frame{
@@
 \frame{
+	\frametitle{Hiccups}
+
+}
+
+\frame{
 \centerline{Thank you!}
+Suggestions and Questions are welcome, suggestions a little more than questions
 }

### 2013-11-05 presentation/ppt.tex
approach again

@@
 
-\section{Introduction}
+\section{Introduction and Related Work}
 
@@
 
-\section{Approach}
+\section{Approach and Progress}
 
 \frame{
-\frametitle{What }
+\frametitle{Our Approach}
 \begin{itemize}
-	\item
+\item What is being done ? 
+	\begin{itemize}
+		\item It's a Replicated state machine (strong or lazy replication) architecture, which had been well studied in Distributed Systems community
+		\item Past work uses DHTs (Onix), Transactional DB (Onix), DFS (Hyperflow)
+	\end{itemize}
+	\item We are trying to study		
+	\begin{itemize}
+		\item Instead of using a general repilication mechanism, can we exploit the fact that network views are being replicated ?
+		\item Could the SDN datapath (replication client) make use of replication timestamps ?
+	\end{itemize}
+
 \end{itemize}
@@
 \frame{
-\frametitle{Our Approach}
+\frametitle{Hierarchical Architecture}
 \begin{itemize}
@@
 	\end{itemize}
-	\item Network Aggregation
+\end{itemize}
+}
+
+\frame{
+\frametitle{Network Aggregation}
 	\begin{itemize}
@@
 		\item Convert flow addtion/removal commands on aggregated network to actual network in an efficient way
-	\end{itemize}
-	\item Timestamp: 
-		\item Handle domain level inconsistencies in "SDN Datapath" (Domain Network)
-		\item Each domain i maintains a vector timestamp T_i[1...n] 
-		\item (Routers in) A Domain may change flow depending on their timestamps
-	\end{itemize}   
-\end{itemize}
-}
 
-\section{Progress}
-
-\frame{
-\frametitle{Network Aggregation}
-\begin{itemize}
-	\item
 	\end{itemize}
@@
 	\begin{itemize}   
-	\item
+		\item Handle domain level inconsistencies in Domain Network
+		\item Each domain i maintains a vector timestamp, conistsing of revision number of domains
+		\item Border routers in a domain may change a flow depending on it's timestamp
 	\end{itemize}   

### 2013-11-05 presentation/ppt.tex
future work, hitches, timestamp, references

@@
 	\begin{itemize}
-		\item Instead of using a general repilication mechanism, can we exploit the fact that network views are being replicated ?
-		\item Could the SDN datapath (replication client) make use of replication timestamps ?
+		\item Instead of just using a general repilication mechanism, can we exploit the fact that network views are being replicated ?
+		\item Could SDN datapath (replication client) make use of replication timestamps ?
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Handle domain level inconsistencies in Domain Network
-		\item Each domain i maintains a vector timestamp, conistsing of revision number of domains
-		\item Border routers in a domain may change a flow depending on it's timestamp
+		\item An example - Link Balancer Controller [4]
+		\item On a flow arriving at an ingress port, a path with minimum maximum utilization is choosen
+		\item Each domain i maintains a vector timestamp, conistsing of revision number of domains' link utilizations, seen so far
+		\item When a flow path is set up at a domain, it is augmented by the timestamp of the domain
+		\item Border routers in a domain may change a flow path if their timestamp is greater than flow's timestamp
+		\item Flow's timestamp is updated as it passes through domains
+		\item what if flow's timestamp is neither greater nor smaller than domain's timestamp ?
 	\end{itemize}   
@@
 \frame{
-\frametitle{Future Work}
+\frametitle{Hitches}
 \begin{itemize}
-	\item 
-	\item Implement using OpenFlow
+	\item Hierarchical Architecture
+	\begin{itemize}
+		\item Suboptimal performance
+	\end{itemize}
+	\item Network Aggregation
+	\begin{itemize}
+		\item May backfire: single internal update may cause multiple updates in aggregated network 
+	\end{itemize}
+	\item Timestamps
+	\begin{itemize}
+		\item Not a true datapath technique
+		\item Improve the performance of control applications, however, does not provide any guarantees
+	\end{itemize}
 \end{itemize}	
@@
 \frame{
-	\frametitle{Hiccups}
-
+\frametitle{Future Work}
+\begin{itemize}
+	\item Understand the trade-off between sub-optimality and frequency of updates, in case of hierarchical architecture
+	\item Revise Network Aggregation data structures so as not to explode network updates
+	\item Implement and simulate each approach on real ISP topology with real traffic data. Analyze performance
+\end{itemize}
 }
@@
 \centerline{Thank you!}
-Suggestions and Questions are welcome, suggestions a little more than questions
+}
+
+\frame{
+\frametitle{References}
+\begin{itemize}
+\item [1] T.
+Koponen, M.
+Casado, N.
+Gude, J.
+Stribling, L.
+Poutievski, M.
+Zhu, R.
+Ramanathan, Y.
+Iwata, H.
+Inoue, T.
+Hama, and S.
+Shenker.
+Onix: a distributed control platform for large-scale production networks.
+In Proceedings of the 9th USENIX OSDI conference , pages 1 - 6, 2010.
+\item [2] A. Tootoonchian and Y. Ganjali. Hyperflow: a distributed control plane for openflow. In Proceedings of the 2010 INM conference, 2010.
+\item [3] Soheil Hassas Yeganeh and Yashar Ganjali.
+Kandoo: a framework for efficient and scalable offloading of control applications.
+In Proc.
+1st Workshop on Hot Topics in Software Defined Networking (HotSDN 2012), pages 19-24, New York, 2012.
+ACM Press.
+\item [4] D. Levin, A. Wundsam, B.Heller, N. Handigol, and A. Feldmann, Logically Centralized State Distribution Trade-offs in Software Defined Networks  in HotSDN, 2012. 
+\end{itemize}
+}
+
+\frame{
+\begin{itemize}
+\frametitle{References}
+\item [5] Stefan Schmid and Jukka Suomela, Exploiting Locality in Distributed SDN Control, in HotSDN, 2013. 
+\item [6] Advait Dixit, Fang Hao, Sarit Mukherjee, T. V. Lakshman and Ramana Kompella, Towards an Elastic Distributed SDN Controller, in HotSDN, 2013.
+\item [7] Nate Foster, Rob Harrison, Matthew L. Meola, Michael J. Freedman, Jennifer Rexford and David Walker, Frenetic: A High-Level Language for OpenFlow Networks. ACM PRESTO 2010.
+\end{itemize}
 }

### 2013-11-05 presentation/ppt.tex
timestamp slide

@@
 		\item On a flow arriving at an ingress port, a path with minimum maximum utilization is choosen
-		\item Each domain i maintains a vector timestamp, conistsing of revision number of domains' link utilizations, seen so far
-		\item When a flow path is set up at a domain, it is augmented by the timestamp of the domain
-		\item Border routers in a domain may change a flow path if their timestamp is greater than flow's timestamp
-		\item Flow's timestamp is updated as it passes through domains
-		\item what if flow's timestamp is neither greater nor smaller than domain's timestamp ?
+		\item Each domain $i$ maintains a vector timestamp, $T_{i}[1...n]$. $T_{i}[j]$ contains the latest revision number of domain $j$ link utilizations, seen by domain $i$
+		\item Each router in a domain is updated with domain's timestamp
+		\item When a flow path is being set up at domain $j$, it is augmented by the $T_{j}$
+		\item When a router in a domain $i$ receives a flow with timestamp $T_{j}$, it may change part of flow's path if $T_{j}[i] < T_{i}[i]$ 
+		\item Flow's timestamp is updated as $max\{T_{j}[k], T_{i}[k]\}$ , $k=1...n$ 
+		\item OK, can we do better ? Yes!, a domain can change the entire flow's path according to timestamps.
 	\end{itemize}   

### 2013-11-05 presentation/ppt.tex
net aggr slide

@@
 	\begin{itemize}		
-		\item Share a condensed view of the network: A Network flow table
-		\item Surpress network events such that aggregated network does not change: Reduce communication overhead
-		\item Convert flow addtion/removal commands on aggregated network to actual network in an efficient way
-
+		\item Definition - Create and publish one flow table (referred to as $nflow$) for a given network.
+(1) Support addition of flows to $nflow$ efficiently.
+(2) Handle intra-link failure efficiently, in such a way that $nflow$  does not change.
+		\item $nflow$ schema $(port in, flow in, port out, flow out, priority, attributes)$
+		\item Data structures\footnote{code @ https://github.com/MugiwaraLuffy/ACN/tree/master/simulator/src/graph}:
+		\begin{itemize}
+			\item Add $external$ $nodes$ to the network graph $G$
+			\item Flow graphs: For each external node, $e$, we create/maintain a directed flow graph, $f_e$ as an overlay graph on $G$, rooted at $e$, which is the source.
+End host and other external nodes acts as sink.
+The flow in $f_e$ are divided at nodes among the incident links according to the flow tables at corresponding routers.
+			\item Other data structures include: Path array, Link array 
+		\end{itemize} 		
 	\end{itemize}

### 2013-11-05 presentation/ppt.tex
Ppt update

@@
 \begin{itemize}
-	\item Hierarchical Architecture
+	\item Controllers' view do not span the whole network, but only the 'underlying network'
+	\item A controller behaves as an openflow switch for parent controller (Network Aggregation) and as a controller for its underlying network
+	\item Controller translates OFP commands and network updates between parent controller and underlying networks
+	\item Control applications run on each controller to set the network parameters in its underlying network
+\end{itemize}
+}
+
+\frame{
+\frametitle{Simulation}
 \begin{itemize}
-		\item Controllers' view do not span the whole network, but only the 'underlying network'
-		\item Each controller behaves as a giant openflow switch for parent controller (Network Aggregation)
-		\item Controller translates OFP commands and network updates between parent controller and underlying networks
-	\end{itemize}
+	\item Wrote a simulator to simulate a hierarchical controller architecture\footnote{code @ https://github.com/MugiwaraLuffy/ACN/tree/master/simulator/src/graph}
+	\item A minimal API for control plane, data plane, controller behavior, etc have been defined
+	\item A $Node$ is an element in the network which implements a control plane interface 
+	\item A $Switch$ extends a $Node$ and contains a data plane
+	\item A $Controller$ extends a $Node$ and implements standard controller API
+	\item $DataNetwork$ contains all the switches which handle actual data packets
+	\item $ControlNetwork$ contains hierarchical layers of controllers
 \end{itemize}

### 2013-11-05 presentation/ppt.tex
Update again

@@
 \begin{itemize}
-	\item Wrote a simulator to simulate a hierarchical controller architecture\footnote{code @ https://github.com/MugiwaraLuffy/ACN/tree/master/simulator/src/graph}
-	\item A minimal API for control plane, data plane, controller behavior, etc have been defined
-	\item A $Node$ is an element in the network which implements a control plane interface 
+	\item Wrote a simulator to simulate a hierarchical controller architecture\footnote{code @ https://github.com/MugiwaraLuffy/ACN/tree/master/simulator/src/swiconsim}
+	\item A minimal set of APIs for control plane, data plane, controller behavior, etc has been defined
+	\item A $Node$ is an element in the management network which implements $IControlPlane$ interface 
 	\item A $Switch$ extends a $Node$ and contains a data plane
-	\item A $Controller$ extends a $Node$ and implements standard controller API
+	\item A $Controller$ extends a $Node$ and implements standard controller APIs
 	\item $DataNetwork$ contains all the switches which handle actual data packets
-	\item $ControlNetwork$ contains hierarchical layers of controllers
+	\item A $Controller$ controls a $Node$ which could be a $Switch$ or another $Controller$
 \end{itemize}

### 2013-11-05 presentation/ppt.tex
spelling errors

@@
 		\item Flat distributed controller architecture
-		\item Appliations are agnostic to underlying state distribution
+		\item Applications are agnostic to underlying state distribution
 		\item Provides a logically centralized global view of the network
@@
 	\begin{itemize}
-		\item Instead of just using a general repilication mechanism, can we exploit the fact that network views are being replicated ?
+		\item Instead of just using a general replication mechanism, can we exploit the fact that network views are being replicated ?
 		\item Could SDN datapath (replication client) make use of replication timestamps ?
@@
 		\item An example - Link Balancer Controller [4]
-		\item On a flow arriving at an ingress port, a path with minimum maximum utilization is choosen
+		\item On a flow arriving at an ingress port, a path with minimum maximum utilization is chosen
 		\item Each domain $i$ maintains a vector timestamp, $T_{i}[1...n]$. $T_{i}[j]$ contains the latest revision number of domain $j$ link utilizations, seen by domain $i$

### 2013-11-20 presentation/ppt.tex
ppt

@@
 \begin{itemize}
+	\item Hierarchical Architecture
+	\begin{itemize}
 		\item Controllers' view do not span the whole network, but only the 'underlying network'
-	\item A controller behaves as an openflow switch for parent controller (Network Aggregation) and as a controller for its underlying network
+		\item Each controller behaves as a giant openflow switch for parent controller (Network Aggregation)
 		\item Controller translates OFP commands and network updates between parent controller and underlying networks
-	\item Control applications run on each controller to set the network parameters in its underlying network
 	\end{itemize}
-}
-
-\frame{
-\frametitle{Simulation}
-\begin{itemize}
-	\item Wrote a simulator to simulate a hierarchical controller architecture\footnote{code @ https://github.com/MugiwaraLuffy/ACN/tree/master/simulator/src/swiconsim}
-	\item A minimal set of APIs for control plane, data plane, controller behavior, etc has been defined
-	\item A $Node$ is an element in the management network which implements $IControlPlane$ interface 
-	\item A $Switch$ extends a $Node$ and contains a data plane
-	\item A $Controller$ extends a $Node$ and implements standard controller APIs
-	\item $DataNetwork$ contains all the switches which handle actual data packets
-	\item A $Controller$ controls a $Node$ which could be a $Switch$ or another $Controller$
 \end{itemize}

### 2013-12-09 paper/paper.tex
paper draft

\title{Hierarchical Controller Architecture in Software Defined Networks}
\author{
        Author 1 \\
        UIUC\\
            \and
        Author 2\\
        UIUC\\
}
\date{\today}

\documentclass[10pt, twocolumn]{article}

\begin{document}
\maketitle
\begin{abstract}
Network control plane, in Software Defined Networks (SDN), provides a global view of physical network to control applications (such as routing), which can be developed as a centralized application rather than a distributed one. Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple nodes called controllers, to achieve desired level of scalability, reliability and responsiveness.

Traditional architecture involves multiple controllers, each operating on a subset of switches and maintaining a global view of entire network. We propose a different where controllers are arranged in a hierarchy and only maintain a part of network view. This improves responsiveness and reduces inconsistency ........(numbers).

Besides, we also propose an optimization in traditional controller architecture. We show 
\end{abstract}

\section{Introduction}
Forwarding or data plane consists of a collection of forwarding elements (switches), connected to form a network, which is inherently distributed in nature. Control plane consists of a collection of control elements, which install forwarding rules on forwarding elements as dictated by control logic or application, running on control elements. Some classes of control application are routing, traffic engineering (TE) and enforcing network policies. In traditional network, there is a tight coupling and one to one mapping between control elements and forwarding elements, with one pair on each switch. Besides, the role of control elements had been trivial, which is to merely install instructions given by control application on corresponding forwarding element as forwarding rules. It is the control application which runs distributed protocol to synchronize the state of forwarding elements. The design of these distributed protocols differs from application to application. For example, link state routing makes switches share their neighbors while distance vector routing makes switches share their routing tables.

The emergence of Software Defined Network (SDN) has induced a separation between forwarding and control elements and has significantly improved the role of control plane. The control element may no longer be present on the switch but move to a separate device called controller, and uses predefined protocol such as OpenFlow [openflow-ref], to install forwarding rules. There may be a one-to-many mapping from controllers to switches. At one extreme, control plane may provide control application with global network view, which can then be implemented as a centralized application, rather than a distributed one. Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple controllers, to achieve desired level of scalability, reliability and responsiveness.

Conventional control plane architecture involves multiple controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application. Instructions (forwarding rules) from control application are sent to target switches directly or through other controllers. Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT). Consequently, this introduces trade-off between network state inconsistency and responsiveness. Results from [logically-centralized-ref] suggests that control state inconsistency significantly degrades performance of logically centralized control applications agnostic to the underlying state distribution.

The state inconsistency problem arises from the fact that each controller try to maintain the complete global view of network. Besides, we believe it is difficult (from scalability point of view) for a controller (a COTS machine) to handle the entire network view. Motivated by these observation, we propose a hierarchical controller architecture. We show ....................

\paragraph{Outline}
The remainder of this article is organized as follows. Section~\ref{sec:hierarchical} presents Hierarchical Controller Architecture. We present the experimental results in section~\ref{sec:eval}. In section~\ref{sec:timestamps}, we present an optimization to conventional controller architecture to help improve the performance of application even if states between controllers is inconsistent. Related work is presented in section~\ref{sec:related}. We discuss ............Our new and exciting results are described in Section~\ref{results}.

\section{Hierarchical Controller Architecture}
\label{sec:hierarchical}
We propose a control plane architecture in which controllers are arranged in the form of a tree over the physical network. Switches form the leaf nodes in the tree, while controllers form the internal nodes. An internal node acts as a switch for it's parent and acts as a controller for it's children. A node aggregate it's children network and converts it to a switch abstraction which is then exposed to the parent controller. This switch abstraction is explained in detail in section~\ref{subsec:aggr}. A controller's view does not span the whole network but only the network of it's children. A node translates instructions (forwarding rules) from it's parent controller into forwarding rules for it's underlying network. Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller. The translation process is explained in the following section.

\subsection{One switch abstraction}
\label{subsec:aggr}

\subsection{An example}
\label{example}
Let's consider an example to understand the benefits of hierarchical architecture. Let there be a TE control application running on $d$ controllers, with each controller operating on $d$ switches. Network state consists of current utilizations of all links. Whenever a new flow arrives, application choses a path with minimum maximum utilization and assigns it to the flow. Each time a path of length $l$ is assigned to the flow, there could be a maximum of $l*d$ link utilization updates and $l$ forwarding rule messages to be sent between controllers. Now, consider an architecture where controllers are arranged in hierarchical manner. There are $d+1$ controllers. The first $d$ controllers (called type 0) maintain $d$ switches each and acts as a switch to the last controller (called type 1), which in turn maintains the other $d$ controllers. Type 0 controllers keep track of link utilizations only for the links that connect it's children switches. Type 1 controller keep track of links that runs from switches belonging to one controller domain to another. Same TE control application runs on each controller. When a new flow arrives, a switch sends $packet_in$ message to it's type 0 controller. Type 0 controller checks if the path of the flow belongs entirely to it's underlying network or not. If yes, it uses TE application to assign the path to the flow and no updates are sent to any controller. However, if the destination is outside it's domain, type 0 controller sends a $packet_in$ message to it's parent controller. Parent controller uses TE control application to assign a paths of type 0 controllers. Forwarding rules are sent to type 0 controllers, which in turn sets up forwarding rules in their network. If the path length is $l$, there will be a maximum of $l+1$ forwarding rule messages between controllers and no link utilization updates. This significant reduction in communication overhead comes at the cost of sub-optimality in the application performance. We show experimentally that the proposed architecture significantly reduces communication overhead while only modestly degrading application performance in the next section.

\section{Evaluation}
\label{sec:eval}
In this section we describe the results.

\section{Timestamps}
\label{sec:timestamps}
We worked hard, and achieved very little.

\section{Discussion}
\label{sec:discuss}

\section{Related Work}
\label{sec:related}

\section{Conclusion}
\label{sec:conc}

\section{Acknowledgements}
\label{sec:ack}

\bibliographystyle{abbrv}
\bibliography{main}

\end{document}
This is never printed

### 2013-12-14 paper/paper.tex
author list changed to length 3
Evaluation sections

@@
         Author 1 \\
-        UIUC\\
             \and
         Author 2\\
-        UIUC\\
+        %UIUC\\
+			\and
+		Author 3\\
 }
@@
 \documentclass[10pt, twocolumn]{article}
+\usepackage{url,graphicx,tabularx,array,geometry, caption, subcaption, float}
+
+%-- Commands for header
+\renewcommand{\title}[1]{\textbf{#1}\\}
+\renewcommand{\line}{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}\hline\\\end{tabularx}\\[-0.10cm]}
+\newcommand{\leftright}[2]{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}#1%
+& #2\\\end{tabularx}\\[-0.10cm]}
+
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 Related work is presented in section~\ref{sec:related}.
-We discuss ............Our new and exciting results are described in Section~\ref{results}.
+We discuss ............Our new and exciting results are described in Section~\ref{sec:eval}.
 
@@
 \label{sec:eval}
-In this section we describe the results.
+We simulated a dummy control application on both hierarchical as well as flat controller architecture, with same network topology and flow arrival distribution.
+The details of simulation, metrics and results are presented in this section.
+
+\subsection{Implementation}
+\label{sec:implementation}
+We wrote our own simulator, which is available at [TODO REF simulator].
+We did not use an existing simulator/controller such as Mininet with Floodlight [TODO ref] for the following two reasons: (1) Existing controller lacks the support for flexible controller architecture and are not suitable for hierarchical arrangement of controllers.
+Modifying existing controllers such as Floodlight to add support for One-switch abstraction (section~\ref{subsec:aggr}), is not a good idea because it requires code/design changes in their primitive modules.
+Besides, we were not looking for a feature rich controller at that point of time.
+(2) Existing simulators do not provide controls to regulate replication between controllers or to measure communication overhead (number of messages exchanged ) between controllers, which we needed for evaluation of control plane architecture.
+Given the above requirements, it seemed easier to write a simulator which provides one-switch abstraction support for the controllers.
+There are three prime interfaces: (1) {\it Switch Data Plane:} It is implemented by a switch to create forwarding elements to handle packet forwarding according to forwarding rules.
+(2) {\it Switch Control Plane:} It is implemented by both controllers and switches and is used to establish connections to parent controllers to send packets which do not match any forwarding rule and to receive forwarding rules from parent controller.
+(3) {\it Controller Control Plane:} It is implemented by controllers to establish connections to children switches/controllers.
+It is used to receive packets which do not match any forwarding rule and to send forwarding rules to children switches/controllers.Simulator defines two networks: (1) {\t Data Network:} This network consists of all switches (and no controller) connected through links.
+(2) {\it Management Network:} This network consists of all controllers and switches connected through Controller - Switch Control Plane interfaces.
+This network allows flexible controller architecture, which could be hierarchical or flat.
+
+To observe the effect of different architecture in real world applications, we used SWIM project [TODO ref swim], which provides synthetic representative test workloads  by sampling historical MapReduce cluster traces using [swim paper ref TODO].
+The original trace comes from historical Hadoop traces on a 600 machine cluster at Facebook and spans 6 months, and contains roughly 1 million jobs.
+The representative trace file contains ~6000 jobs.
+The trace was simulated with a cluster of 600 machines (as in the original cluster) connected through a network of 200 switches and 10 Gbps links.
+The cluster is divided into 10 control domains, with 60 hosts, 20 switches and one controller for each domain.
+Each switch in a domain is connected to 1 to 4 switches within the same domain, with degree being chosen randomly.
+Each domain contains two Top-of-the-rack switches which inter-connects domains.
+For hierarchical architecture, there are 11 controllers: 1 controller per domain and one root controllers which manages the other 10 controllers.
+While flat architecture consists of 10 peer controllers, one or each domain.
+
+Control application is a TE-cum-routing application.
+When a new flow arrives, control application finds the path with minimum maximum utilization and sets up forwarding rules (based on IP destination and incoming ports) along the path.
+If multiple such paths exists (which is frequently the case, due to big network), shortest path (in term of number of hops) is chosen.
+This application runs on each controller in both hierarchical and flat architecture.
+However, in the hierarchical architecture, the view of application is limited and may consists of controllers or switches as forwarding elements, while in the flat architecture, control application is presented with global view of the network.
+For simplicity, we have kept network topology as static i.e.
+links/nodes do not fail and/or created during the simulation.
+For each Hadoop job, a dummy MapReduce application choses directory, map and reduce nodes randomly in the network, thus, creating a number of flows which are handled by Control application running on controllers.
+The duration of flow depends on the bytes needed to transfer and assigned rates.
+
+\subsection{Results}
+\label{subsec:results}
+
+\begin{figure}[t]
+\centering
+\begin{subfigure}{.5\textwidth}
+\label{fig:path}
+\includegraphics[scale=0.3]{path}
+\caption{Application performance: path length for flows}
+\end{subfigure}
+
+\begin{subfigure}{.5\textwidth}
+\label{fig:util}
+\includegraphics[scale=0.3]{util}
+\caption{Application performance: Maximum link utilization for flows}
+\end{subfigure}
+\caption{Application performance}
+\end{figure}
+
+\begin{figure}
+\label{fig:msg}
+\includegraphics[scale=0.3]{msg}
+\caption{Communication overhead per flow}
+\end{figure}
+
+We simulated the trace file for both flat and hierarchical architecture, with same topology and same assignment of task nodes for each job.
+We calculated two metrics: (1) The performance of control application which is measured as the length path (number of hops) assigned to flows (Figure ~\ref{fig:path}) and maximum link utilization encountered by the flow (Figure ~\ref{fig:util}).
+While the flat architecture gives optimal results, hierarchical architecture provides sub-optimal results.
+(2) Communication overhead during flow set up, which is measured in term of number of messages passed between switches and controllers as well as within controllers (Figure ~\ref{fig:msg}).
+The simulation assumes instant replication of network state between controllers.
+Results show that hierarchical architecture reduces the communication overhead between controllers by an order of magnitude while modestly deviating from optimal solution.
+
+\begin{table*}
+\centering
+\begin{tabular}{| L{5cm} | L{3cm} | L{3cm} | L{5cm} |}
+\hline
+& {Average path length} & {Average maximum utilization} & {Average communication overhead per flow} \\ \hline
+Flat architecture & 7.81 & 24.62 & 153.54 \\ \hline
+Hierarchical architecture & 9.17 & 27.78 & 14.52 \\ \hline
+\end{tabular}
+\end{table*}%
 
@@
 \end{document}
-This is never printed

### 2013-12-14 simulator/src/swiconsim/SwiConSim.java
everything

@@ -1,6 +1,7 @@
 package swiconsim;
 
 import java.util.ArrayList;
+import java.util.Iterator;
 import java.util.List;
 
 import swiconsim.controller.Controller;
@@ -20,7 +21,145 @@ import swiconsim.util.PortUtil;
 public class SwiConSim {
 
 	public static void main(String[] args) {
-		samplerun2();
+		samplerun3();
+	}
+	
+	static void samplerun4(){
+		
+	}
+	
+	static void samplerun3(){
+		int nports=10;
+		Controller c = new Controller(1000);
+		
+		Controller c1 = new Controller(1001,c.getId());
+		Switch sw11 = new Switch(11, nports, c1.getId());
+		Switch sw12 = new Switch(12, nports, c1.getId());
+		Switch sw13 = new Switch(13, nports, c1.getId());
+		Switch sw14 = new Switch(14, nports, c1.getId());
+		DataNetwork.getInstance().addEdge(1100001, 1200001, 100);
+		DataNetwork.getInstance().addEdge(1200002, 1300001, 100);
+		DataNetwork.getInstance().addEdge(1300002, 1400001, 100);
+		DataNetwork.getInstance().addEdge(1400002, 1100002, 100);
+		
+		Controller c2 = new Controller(1002,c.getId());
+		Switch sw21 = new Switch(21, nports, c2.getId());
+		Switch sw22 = new Switch(22, nports, c2.getId());
+		Switch sw23 = new Switch(23, nports, c2.getId());
+		Switch sw24 = new Switch(24, nports, c2.getId());
+		DataNetwork.getInstance().addEdge(2100001, 2200001, 100);
+		DataNetwork.getInstance().addEdge(2200002, 2300001, 100);
+		DataNetwork.getInstance().addEdge(2300002, 2400001, 100);
+		DataNetwork.getInstance().addEdge(2400002, 2100002, 100);
+		
+		Controller c3 = new Controller(1003,c.getId());
+		Switch sw31 = new Switch(31, nports, c3.getId());
+		Switch sw32 = new Switch(32, nports, c3.getId());
+		Switch sw33 = new Switch(33, nports, c3.getId());
+		Switch sw34 = new Switch(34, nports, c3.getId());
+		DataNetwork.getInstance().addEdge(3100001, 3200001, 100);
[…]

### 2013-12-14 simulator/src/swiconsim/api/IControlPlane.java

@@ -25,7 +25,7 @@ public interface IControlPlane {
 	 * 
 	 * @param flow
 	 */
-	void addFlow(Flow flow);
+	void addFlow(Flow flow, Packet packet);
 
 	/**
 	 * remove
@@ -43,7 +43,7 @@ public interface IControlPlane {
 	 * @param portNum
 	 * @param port
 	 */
-	void addPort(short portNum, Port port);
+	void addPort(long portNum, Port port);
 
 	/**
 	 * forward an in-pkt to controller

### 2013-12-14 simulator/src/swiconsim/api/IControllerSouthBound.java

@@ -1,6 +1,7 @@
 package swiconsim.api;
 
 import swiconsim.messages.Message;
+import swiconsim.packet.Packet;
 import swiconsim.flow.Flow;
 
 /**
@@ -18,7 +19,7 @@ public interface IControllerSouthBound {
 	 * @param swid
 	 * @param flow
 	 */
-	void addFlowToSwitch(long swid, Flow flow);
+	void addFlowToSwitch(long swid, Flow flow, Packet p);
 	/**
 	 * @param swid
 	 * @param flow

### 2013-12-14 simulator/src/swiconsim/api/IDataNetwork.java

@@ -2,6 +2,7 @@ package swiconsim.api;
 
 import java.util.Map;
 
+import swiconsim.link.Link;
 import swiconsim.node.Node;
 import swiconsim.packet.Packet;
 
@@ -30,7 +31,7 @@ public interface IDataNetwork {
 	 * @param swid2
 	 * @param portNum2
 	 */
-	void addEdge(long swid1, short portNum1, long swid2, short portNum2);
+	void addEdge(long swid1, short portNum1, long swid2, short portNum2, int capacity);
 
 	/**
 	 * add an edge by portid
@@ -38,7 +39,7 @@ public interface IDataNetwork {
 	 * @param portId1
 	 * @param portId2
 	 */
-	void addEdge(long portId1, long portId2);
+	void addEdge(long portId1, long portId2, int capacity);
 
 	/**
 	 * pkt coming into data network through portid
@@ -64,7 +65,7 @@ public interface IDataNetwork {
 	/**
 	 * @return
 	 */
-	Map<Long, Long> getLinks();
+	Map<Long, Link> getLinks();
 	
 
 }

### 2013-12-14 simulator/src/swiconsim/api/IManagementNetwork.java

@@ -1,6 +1,8 @@
 package swiconsim.api;
 
 import swiconsim.messages.Message;
+import swiconsim.node.ManagementNode;
+
 import swiconsim.controller.Controller;
 import swiconsim.node.Node;
 
@@ -24,7 +26,7 @@ public interface IManagementNetwork {
 	 * @param id
 	 * @param cont
 	 */
-	void registerController(long id, Controller cont);
+	void registerController(long id, ManagementNode cont);
 
 	/**
 	 * notify controller

### 2013-12-14 simulator/src/swiconsim/api/ISwitchDataPlane.java

@@ -13,7 +13,7 @@ public interface ISwitchDataPlane {
 	 * @param in_port
 	 * @return
 	 */
-	boolean receivePkt(Packet pkt, short in_port);
+	boolean receivePkt(Packet pkt, long in_port);
 	/**
 	 * @param pkt
 	 * @param out_port
@@ -24,5 +24,5 @@ public interface ISwitchDataPlane {
 	 * @param pkt
 	 * @param out_port
 	 */
-	void sendPkt(Packet pkt, short out_port);
+	void sendPkt(Packet pkt, long out_port);
 }

### 2013-12-14 simulator/src/swiconsim/controller/Controller1.java

@@ -29,7 +29,6 @@ import swiconsim.nwswitch.port.Port;
 import swiconsim.packet.Packet;
 import swiconsim.util.IPUtil;
 import swiconsim.util.PortUtil;
-import swiconsim.node.ManagementNode;
 
 /**
  * @author praveen
@@ -37,7 +36,7 @@ import swiconsim.node.ManagementNode;
  *         Controller
  * 
  */
-public class Controller extends ManagementNode implements IControlPlane, IController,
+public class Controller1 extends Controller implements IControlPlane, IController,
 		IControllerSouthBound {
 	private static Logger logger = Logger.getLogger("sim:");
 	public List<Long> nodes;
@@ -46,14 +45,14 @@ public class Controller extends ManagementNode implements IControlPlane, IContro
 	public ControllerSouthBound csb;
 	public long parent=0;
 
-	public Controller(long id, long cid) {
+	public Controller1(long id, long cid) {
 		this(id);
 		DataNetwork.getInstance().registerNode(id, this);
 		registerWithController(cid);
 		parent=cid;
 	}
 
-	public Controller(long id) {
+	public Controller1(long id) {
 		super(id);
 		this.id = id;
 		this.ports = new TreeMap<Long, Port>();
@@ -128,7 +127,6 @@ public class Controller extends ManagementNode implements IControlPlane, IContro
 		//System.out.println("BFS on " + in_port + "->" + out_port);
 		ArrayBlockingQueue<Long> queue = new ArrayBlockingQueue<Long>(nodes.size());
 		ArrayList<Long> closed = new ArrayList<Long>(); 
-		ArrayList<Long> open = new ArrayList<Long>();
 		HashMap<Long, Long> parent = new HashMap<Long, Long>();
 		ArrayList<Long> path = new ArrayList<Long>();
 		queue.add(in_port);
@@ -140,7 +138,6 @@ public class Controller extends ManagementNode implements IControlPlane, IContro
 			queue.poll();
 			
 			closed.add(reverseLookup(port));
-			open.remove(reverseLookup(port));
 			//System.out.println("BFS: switch " + port);
 			//path.add(sw);
 			
@@ -179,13 +176,12 @@ public class Controller extends ManagementNode implements IControlPlane, IContro
 				long port3 = DataNetwork.getInstance().links.get(swPort.getId()).port2;
 				long sw2 = reverseLookup(port3);
 				if(!nodes.contains(sw2)) continue;
[…]

### 2013-12-14 simulator/src/swiconsim/controller/Controller2.java

@@ -16,6 +16,7 @@ import java.util.logging.Logger;
 
 import swiconsim.link.Link;
 import swiconsim.messages.Message;
+import swiconsim.messages.MessageType;
 import swiconsim.api.IController;
 import swiconsim.api.IControllerSouthBound;
 import swiconsim.api.IControlPlane;
@@ -29,6 +30,7 @@ import swiconsim.nwswitch.port.Port;
 import swiconsim.packet.Packet;
 import swiconsim.util.IPUtil;
 import swiconsim.util.PortUtil;
+import swiconsim.util.SwitchFlowPair;
 import swiconsim.node.ManagementNode;
 
 /**
@@ -37,28 +39,20 @@ import swiconsim.node.ManagementNode;
  *         Controller
  * 
  */
-public class Controller extends ManagementNode implements IControlPlane, IController,
-		IControllerSouthBound {
+public class Controller2 extends ManagementNode implements IControlPlane, IController,
+IControllerSouthBound {
 	private static Logger logger = Logger.getLogger("sim:");
 	public List<Long> nodes;
-	public List<Long> ExternalPorts;
 	public HashMap<PortPair, ArrayList<Long>> paths = new HashMap<PortPair, ArrayList<Long>>();
 	public ControllerSouthBound csb;
-	public long parent=0;
 
-	public Controller(long id, long cid) {
-		this(id);
-		DataNetwork.getInstance().registerNode(id, this);
-		registerWithController(cid);
-		parent=cid;
-	}
 
-	public Controller(long id) {
+	public Controller2(long id) {
 		super(id);
+		//DataNetwork.getInstance().registerNode(id, this);
 		this.id = id;
 		this.ports = new TreeMap<Long, Port>();
 		nodes = new ArrayList<Long>();
-		ExternalPorts = new ArrayList<Long>();
 		csb = new ControllerSouthBound(id, nodes, this);
 		registerWithMgmtNetAsController();
 	}
@@ -76,27 +70,21 @@ public class Controller extends ManagementNode implements IControlPlane, IContro
 			System.out.println("WARNING: No in_port specified for flow " + flow.toString() + " at switch " +id);
 			return;
 		}
-		if(!ExternalPorts.contains(in_port)){
-			System.out.println("WARNING: " + in_port + " does not belong to controllers' " + id + " external ports");
-			return;
-		}
[…]

### 2013-12-14 simulator/src/swiconsim/controller/ControllerSouthBound.java

@@ -1,6 +1,8 @@
 package swiconsim.controller;
 
 import java.util.List;
+import swiconsim.node.ManagementNode;
+
 import java.util.Set;
 import java.util.logging.Logger;
 
@@ -8,24 +10,27 @@ import swiconsim.api.IControllerSouthBound;
 import swiconsim.flow.Flow;
 import swiconsim.messages.Message;
 import swiconsim.messages.MessageType;
+import swiconsim.network.DataNetwork;
 import swiconsim.network.ManagementNetwork;
 import swiconsim.nwswitch.Switch;
 import swiconsim.nwswitch.port.Port;
+import swiconsim.packet.Packet;
 import swiconsim.util.PortUtil;
 
 public class ControllerSouthBound implements IControllerSouthBound {
 	private static Logger logger = Logger.getLogger("sim:");
-	List<Long> switches;
+	public List<Long> switches;
 	long id;
-	Controller controller;
+	ManagementNode controller;
 
 	public ControllerSouthBound(long id, List<Long> switches,
-			Controller controller) {
+			ManagementNode controller) {
 		super();
 		this.id = id;
 		this.switches = switches;
 		this.controller = controller;
 	}
+
 
 	/*
 	 * (non-Javadoc)
@@ -40,7 +45,20 @@ public class ControllerSouthBound implements IControllerSouthBound {
 		case HELLO:
 			long switchId = (Long) msg.getPayload();
 			logger.info(this.id + "Hello from " + switchId);
+			
 			switches.add(switchId);
+			
+			
+			List<Port> swPorts = ManagementNetwork.getInstance()
+					.getNode(switchId).getPorts();
+			for (Port swPort : swPorts) {								
+				controller.addPort(swPort.getId(), swPort);				
+			}
+			
+			break;
+		case PKT_IN:
+			//System.out.println("Controller " + id + " received a packet in. Will do something :)");
+			controller.handlePacket((Packet) msg.getPayload());
 			break;
 		default:
 			break;
@@ -54,8 +72,8 @@ public class ControllerSouthBound implements IControllerSouthBound {
 	 * swiconsim.flow.Flow)
 	 */
 	@Override
-	public void addFlowToSwitch(long swid, Flow flow) {
-		Message msg = new Message(swid, MessageType.OFPFC_ADD, flow, this.id);
+	public void addFlowToSwitch(long swid, Flow flow, Packet packet) {
[…]

### 2013-12-14 simulator/src/swiconsim/controller/PortPair.java

@@ -1,0 +1,6 @@
+package swiconsim.controller;
+
+public class PortPair {
+	long port1;
+	long port2;
+}

### 2013-12-14 simulator/src/swiconsim/controller/Topology.java

@@ -6,6 +6,7 @@ import java.util.Map;
 import java.util.Set;
 
 import swiconsim.host.Host;
+import swiconsim.link.Link;
 import swiconsim.node.Node;
 import swiconsim.nwswitch.Switch;
 import swiconsim.nwswitch.port.Port;
@@ -17,10 +18,10 @@ import swiconsim.nwswitch.port.Port;
  */
 public class Topology {
 	HashMap<Node, List<Port>> nodePorts;
-	Map<Long, Long> edges;
+	Map<Long, Link> edges;
 	Set<Host> hosts;
 
-	public Topology(HashMap<Node, List<Port>> nodePorts, Map<Long, Long> edges, Set<Host> hosts) {
+	public Topology(HashMap<Node, List<Port>> nodePorts, Map<Long, Link> edges, Set<Host> hosts) {
 		super();
 		this.nodePorts = nodePorts;
 		this.edges = edges;

### 2013-12-14 simulator/src/swiconsim/flow/Action.java

@@ -7,8 +7,8 @@ package swiconsim.flow;
  */
 public class Action {
 	ActionType type;
-	int value;
-	public Action(ActionType type, int value) {
+	long value;
+	public Action(ActionType type, long value) {
 		super();
 		this.type = type;
 		this.value = value;
@@ -19,10 +19,10 @@ public class Action {
 	public void setType(ActionType type) {
 		this.type = type;
 	}
-	public int getValue() {
+	public long getValue() {
 		return value;
 	}
-	public void setValue(int value) {
+	public void setValue(long value) {
 		this.value = value;
 	}
 	

### 2013-12-14 simulator/src/swiconsim/flow/Match.java

@@ -11,7 +11,7 @@ import swiconsim.util.IPUtil;
  */
 public class Match {
 	/** Input switch port */
-    public short in_port = 0;
+    public long in_port = 0;
 
     /** IP source address */
     public int nw_src = 0;
@@ -32,7 +32,7 @@ public class Match {
      * @param nw_dst
      *  
      */
-	public Match(short in_port, int nw_src, int nw_dst) {
+	public Match(long in_port, int nw_src, int nw_dst) {
 		super();
 		this.in_port = in_port;
 		this.nw_src = nw_src;
@@ -45,7 +45,7 @@ public class Match {
 		super();
 		switch(field){
 		case IN_PORT:
-			this.in_port = (short) value;
+			this.in_port = (long) value;
 			break;
 		case SRC:
 			this.nw_src = value;
@@ -60,7 +60,7 @@ public class Match {
 		}
 	}
 
-	public Match(short in_port, int nw_src, int nw_src_mask, int nw_dst,
+	public Match(long in_port, int nw_src, int nw_src_mask, int nw_dst,
 			int nw_dst_mask) {
 		super();
 		this.in_port = in_port;
@@ -70,11 +70,19 @@ public class Match {
 		this.nw_dst_mask = nw_dst_mask;
 	}
 
-	public short getIn_port() {
+	public Match(Match match) {
+		this.in_port = match.in_port;
+		this.nw_src = match.nw_src;
+		this.nw_src_mask = match.nw_src_mask;
+		this.nw_dst = match.nw_dst;
+		this.nw_dst_mask = match.nw_dst_mask;
+	}
+
+	public long getIn_port() {
 		return in_port;
 	}
 
-	public void setIn_port(short in_port) {
+	public void setIn_port(long in_port) {
 		this.in_port = in_port;
 	}
 
@@ -132,8 +140,9 @@ public class Match {
 	}
 
 	public boolean isMatch(PacketIdentifier pktIden){
-		short pktin_port = this.in_port==0?0:pktIden.getIn_port();
+		long pktin_port = this.in_port==0?0:pktIden.getIn_port();
 		Match candidate = new Match(pktin_port, pktIden.getNw_src(),  this.nw_src_mask, pktIden.getNw_dst(), this.nw_dst_mask);
+		//System.out.println("Canddate: " + candidate + ". This: " + toString());
 		if(this.toString().equals(candidate.toString())) return true;
 		return false;
 	}

### 2013-12-14 simulator/src/swiconsim/host/Host.java

@@ -2,6 +2,8 @@ package swiconsim.host;
 
 import java.util.logging.Logger;
 
+import swiconsim.network.DataNetwork;
+import swiconsim.network.ManagementNetwork;
 import swiconsim.nwswitch.port.Port;
 import swiconsim.packet.Packet;
 import swiconsim.util.PortUtil;
@@ -53,13 +55,22 @@ public class Host {
 
 	public void receivePkt(Packet pkt) {
 		logger.info("host pkt rcvd: " + pkt.toString());
+		if(pkt.last==0){
+			DataNetwork.getInstance().HopCount.put(pkt, pkt.nhops);
+			DataNetwork.getInstance().UtilizationTracker.put(pkt, pkt.maxutil);
+		}else{
+			//System.out.println("Got last packet. Not updating hop count");
+		}
+		if(!ManagementNetwork.getInstance().MessageCount.containsKey(pkt)){
+			ManagementNetwork.getInstance().MessageCount.put(pkt, 0);
+		}					
 		rx++;
 	}
 
 	public void sendPkt(Packet pkt) {
 		logger.info("host pkt sent: " + pkt.toString());
 		this.port.getSw().receivePkt(pkt,
-				PortUtil.getPortNumFromPortId(this.port.getId()));
+				this.port.getId());
 		tx++;
 	}
 
@@ -67,5 +78,23 @@ public class Host {
 	public String toString() {
 		return "Host [ip=" + ip + ", id=" + id + ", rx=" + rx + ", tx=" + tx
 				+ ", port=" + port.getId() + "]";
+	}
+
+	public void startFlow(Packet pkt) {
+		// TODO Auto-generated method stub
+		logger.info("host starting flow: " + pkt.toString());
+		this.port.getSw().receivePkt(pkt,
+				this.port.getId());
+		tx++;
+		
+	}
+
+	public void endFlow(Packet pkt) {
+		// TODO Auto-generated method stub
+		logger.info("host terminating flow: " + pkt.toString());
+		this.port.getSw().receivePkt(pkt,
+				this.port.getId());
+		tx++;
+
 	}
 }

### 2013-12-14 simulator/src/swiconsim/link/Link.java

@@ -1,0 +1,26 @@
+package swiconsim.link;
+
+public class Link {
+	public long port1;
+	public long port2;
+	public int capacity;
+	public double utilization;
+	
+	public Link(long port1, long port2, int capacity, int utilization){
+		this.port1 = port1;
+		this.port2 = port2;
+		this.capacity = capacity;
+		this.utilization = utilization;
+	}
+	
+	public Link(Link other){
+		this.port1 = other.port1;
+		this.port2 = other.port2;
+		this.capacity = other.capacity;
+		this.utilization = other.utilization;
+	}
+	
+	public String toString(){
+		return ""+port2;
+	}
+}

### 2013-12-14 simulator/src/swiconsim/messages/Message.java

@@ -1,5 +1,7 @@
 package swiconsim.messages;
 
+import swiconsim.packet.Packet;
+
 
 
 /**
@@ -12,6 +14,7 @@ public class Message {
 	MessageType type;
 	Object payload;
 	long from;
+	public Packet packet;
 	
 	public long getTo() {
 		return to;
@@ -29,12 +32,13 @@ public class Message {
 		this.from = from;
 	}
 
-	public Message(Long to, MessageType type, Object payload, Long from) {
+	public Message(Long to, MessageType type, Object payload, Long from, Packet p) {
 		super();
 		this.to = to;
 		this.type = type;
 		this.payload = payload;
 		this.from = from;
+		this.packet = p;
 	}
 
 	public MessageType getType() {

### 2013-12-14 simulator/src/swiconsim/messages/MessageType.java

@@ -5,5 +5,7 @@ public enum MessageType {
 			// switch
 	PKT_IN, // switch forwarding an in-pkt to controller
 	OFPFC_ADD, // Add a flow on a switch
-	OFPFC_DELETE // delete a flow from a switch
+	OFPFC_DELETE, // delete a flow from a switch
+	PEER_ADD, //Message sent by controller to another controller, requesting to add the flow on it's switches
+	PEER_UPDATE
 }

### 2013-12-14 simulator/src/swiconsim/network/DataNetwork.java

@@ -1,10 +1,13 @@
 package swiconsim.network;
 
 import java.util.HashMap;
+
+import java.util.Iterator;
 import java.util.Map;
 import java.util.logging.Logger;
 
 import swiconsim.api.IDataNetwork;
+import swiconsim.link.Link;
 import swiconsim.node.Node;
 import swiconsim.nwswitch.Switch;
 import swiconsim.packet.Packet;
@@ -18,20 +21,22 @@ import swiconsim.util.PortUtil;
 public class DataNetwork implements IDataNetwork {
 	private static DataNetwork instance = null;
 	Map<Long, Node> nodeMap;
-	Map<Long, Long> links;
+	public Map<Long, Link> links;
 	private static Logger logger = Logger.getLogger("sim:");
+	public Map<Packet, Integer> HopCount = new HashMap<Packet, Integer>();
+	public Map<Packet, Double> UtilizationTracker = new HashMap<Packet, Double>();
 
 	public Map<Long, Node> getNodeMap() {
 		return nodeMap;
 	}
 
-	public Map<Long, Long> getLinks() {
+	public Map<Long, Link> getLinks() {
 		return links;
 	}
 
 	protected DataNetwork() {
 		nodeMap = new HashMap<Long, Node>();
-		links = new HashMap<Long, Long>();
+		links = new HashMap<Long, Link>();
 	}
 
 	public static DataNetwork getInstance() {
@@ -41,6 +46,20 @@ public class DataNetwork implements IDataNetwork {
 		return instance;
 	}
 
+	public boolean isLink(long portid1, long portid2){
+		if(this.links.containsKey(portid1) && this.links.get(portid1).port2==portid2){
+			return true;
+		}
+		return false;
+		//return false;
+	}
+	
+	public long getOtherPort(long portId){
+		if(this.links.containsKey(portId)){
+			return this.links.get(portId).port2;
+		}
+		return -1;
+	}
 	/*
 	 * (non-Javadoc)
 	 * 
@@ -49,7 +68,9 @@ public class DataNetwork implements IDataNetwork {
 	 */
 	@Override
 	public void registerNode(long id, Node node) {
+		//System.out.println("Registering node " + id);
 		this.nodeMap.put(id, node);
+		//System.out.println(this.nodeMap.keySet());
 	}
 
 	/*
@@ -58,9 +79,9 @@ public class DataNetwork implements IDataNetwork {
[…]

### 2013-12-14 simulator/src/swiconsim/network/ManagementNetwork.java

@@ -1,10 +1,14 @@
 package swiconsim.network;
 
 import java.util.HashMap;
+import swiconsim.node.ManagementNode;
+
 import java.util.Map;
 import java.util.logging.Logger;
 
+import swiconsim.link.Link;
 import swiconsim.messages.Message;
+import swiconsim.messages.MessageType;
 import swiconsim.api.IDataNetwork;
 import swiconsim.api.IManagementNetwork;
 import swiconsim.controller.Controller;
@@ -23,10 +27,14 @@ import swiconsim.util.PortUtil;
 public class ManagementNetwork implements IManagementNetwork, IDataNetwork {
 	private static ManagementNetwork instance = null;
 	Map<Long, Node> nodeMap;
-	Map<Long, Controller> mgmtContMap;
+	Map<Long, ManagementNode> mgmtContMap;
 	private static Logger logger = Logger.getLogger("sim:");
+	public int SwitchToControllerCount=0;
+	public int ControllerToSwitchCount=0;
+	public int ControllerToControllerCount=0;
+	public HashMap<Packet, Integer> MessageCount = new HashMap<Packet, Integer>();
 
-	Map<Long, Long> links;
+	Map<Long, Link> links;
 	Map<Long, Long> virtualPortIdMap;
 	
 	public Map<Long, Long> getVirtualPortIdMap() {
@@ -35,8 +43,8 @@ public class ManagementNetwork implements IManagementNetwork, IDataNetwork {
 
 	protected ManagementNetwork() {
 		nodeMap = new HashMap<Long, Node>();
-		mgmtContMap = new HashMap<Long, Controller>();
-		links = new HashMap<Long, Long>();
+		mgmtContMap = new HashMap<Long, ManagementNode>();
+		links = new HashMap<Long, Link>();
 		virtualPortIdMap = new HashMap<Long, Long>();
 	}
 
@@ -67,7 +75,7 @@ public class ManagementNetwork implements IManagementNetwork, IDataNetwork {
 	 * swiconsim.controller.Controller)
 	 */
 	@Override
-	public void registerController(long id, Controller cont) {
+	public void registerController(long id, ManagementNode cont) {
 		logger.info("Registering controller : " + id);
 		this.mgmtContMap.put(id, cont);
 	}
@@ -84,6 +92,15 @@ public class ManagementNetwork implements IManagementNetwork, IDataNetwork {
[…]

### 2013-12-14 simulator/src/swiconsim/node/ManagementNode.java

@@ -1,0 +1,24 @@
+package swiconsim.node;
+
+import swiconsim.messages.Message;
+import swiconsim.packet.Packet;
+
+public abstract class ManagementNode extends Node {
+	
+	public ManagementNode(long id){
+		super(id);
+	}
+
+	public void handlePacket(Packet payload) {
+		
+	}
+
+	public void receiveNotificationFromSwitch(Message msg) {
+		
+	}
+
+	public void updateUtilization(Packet pkt) {
+		
+	}
+
+}

### 2013-12-14 simulator/src/swiconsim/node/Node.java

@@ -3,6 +3,7 @@ package swiconsim.node;
 import java.util.ArrayList;
 import java.util.HashMap;
 import java.util.HashSet;
+import java.util.Iterator;
 import java.util.List;
 import java.util.Set;
 import java.util.TreeMap;
@@ -16,18 +17,20 @@ import swiconsim.network.ManagementNetwork;
 import swiconsim.nwswitch.FlowTable;
 import swiconsim.nwswitch.port.Port;
 import swiconsim.packet.Packet;
+import swiconsim.util.IPUtil;
 
 public abstract class Node implements IControlPlane {
 	protected long id;
-	protected TreeMap<Short, Port> ports;
+	public TreeMap<Long, Port> ports;
 	private FlowTable flowTable;
 	protected long cid;
 	private static Logger logger = Logger.getLogger("sim:");
+	public long parent=0;
 
 	public Node(long id) {
 		super();
 		this.id = id;
-		ports = new TreeMap<Short, Port>();
+		ports = new TreeMap<Long, Port>();
 		flowTable = new FlowTable();
 		registerWithMgmtNodeAsNode();
 	}
@@ -43,28 +46,46 @@ public abstract class Node implements IControlPlane {
 	@Override
 	public List<Port> getPorts() {
 		List<Port> ports = new ArrayList<Port>();
-		for (short i = 1; i <= this.ports.size(); i++) {
-			ports.add(this.ports.get(i));
-			if (!ManagementNetwork.getInstance().getVirtualPortIdMap()
+		Iterator<Port> it = this.ports.values().iterator();
+		while(it.hasNext()){
+			ports.add(it.next());
+			/*if (!ManagementNetwork.getInstance().getVirtualPortIdMap()
 					.containsKey(this.ports.get(i).getId())) {
 				ManagementNetwork
 						.getInstance()
 						.getVirtualPortIdMap()
 						.put(this.ports.get(i).getId(),
 								this.ports.get(i).getId());
-			}
+			}*/
+		}
+		return ports;
+	}
+	
+	public List<Port> getAllPorts() {
+		List<Port> ports = new ArrayList<Port>();
+		Iterator<Port> it = this.ports.values().iterator();
+		while(it.hasNext()){
+			ports.add(it.next());
+			/*if (!ManagementNetwork.getInstance().getVirtualPortIdMap()
+					.containsKey(this.ports.get(i).getId())) {
+				ManagementNetwork
+						.getInstance()
[…]

### 2013-12-14 simulator/src/swiconsim/nwswitch/FlowTable.java

@@ -29,23 +29,24 @@ public class FlowTable {
 	/**
 	 * @param flowEntries
 	 */
-	public FlowTable(Map<Flow, FlowCounter> flowEntries) {
+	/*public FlowTable(Map<Flow, FlowCounter> flowEntries) {
 		super();
 		this.flowEntries = flowEntries;
-	}
+	}*/
 
 	public Map<Flow, FlowCounter> getFlowEntries() {
 		return flowEntries;
 	}
 
-	public void setFlowEntries(Map<Flow, FlowCounter> flowEntries) {
+	/*public void setFlowEntries(Map<Flow, FlowCounter> flowEntries) {
 		this.flowEntries = flowEntries;
-	}
+	}*/
 
 	/**
 	 * @param flow
 	 */
 	public void addFlowEntry(Flow flow) {
+		//System.out.println("Adding flow " + flow.toString());
 		this.flowEntries.put(flow, new FlowCounter());
 	}
 

### 2013-12-14 simulator/src/swiconsim/nwswitch/Switch.java

@@ -1,6 +1,7 @@
 package swiconsim.nwswitch;
 
 import java.util.HashMap;
+import java.util.Iterator;
 import java.util.TreeMap;
 
 import swiconsim.api.IControlPlane;
@@ -12,6 +13,7 @@ import swiconsim.node.Node;
 import swiconsim.nwswitch.port.Port;
 import swiconsim.nwswitch.port.PortStatus;
 import swiconsim.packet.Packet;
+import swiconsim.util.IPUtil;
 import swiconsim.util.PortUtil;
 
 /**
@@ -19,10 +21,9 @@ import swiconsim.util.PortUtil;
  * 
  *         Switch - control plane and data plane
  */
-public class Switch extends Node implements IControlPlane, ISwitchDataPlane {
+public class Switch extends Node implements IControlPlane, ISwitchDataPlane, Comparable {
 
 	SwitchDataPlane dp;
-	
 	private FlowTable flowTable;
 
 	public Switch(long id) {
@@ -31,16 +32,20 @@ public class Switch extends Node implements IControlPlane, ISwitchDataPlane {
 
 	public Switch(long id, int numPorts, long cid) {
 		this(id, numPorts);
+		this.parent=cid;
+		//DataNetwork.getInstance().registerNode(id, this);
 		registerWithController(cid);
 	}
 
 	public Switch(long id, int numPorts) {
 		super(id);
 		flowTable = new FlowTable();
-		ports = new TreeMap<Short, Port>();
+		ports = new TreeMap<Long, Port>();
 		for (short i = 1; i <= numPorts; i++) {
-			Port port = new Port(PortUtil.calculatePortId(id, i), PortStatus.UP, this);
-			ports.put(i, port);
+			long portid = PortUtil.calculatePortId(id, i);
+			//System.out.println("Switch " + id + ": Adding port " + portid);
+			Port port = new Port(portid, PortStatus.UP, this);
+			ports.put(portid, port);
 		}
 		
 		dp = new SwitchDataPlane(id, ports, flowTable);
@@ -53,8 +58,11 @@ public class Switch extends Node implements IControlPlane, ISwitchDataPlane {
 	}
 
 	@Override
-	public void addFlow(Flow f) {
+	public void addFlow(Flow f, Packet pkt) {
+		//System.out.println(id + ": Adding flow " + f.toString());
 		flowTable.addFlowEntry(f);
+		//System.out.println(flowTable.toString());
+
 	}
 
 	@Override
[…]

### 2013-12-14 simulator/src/swiconsim/nwswitch/SwitchDataPlane.java

@@ -22,11 +22,11 @@ import swiconsim.packet.PacketIdentifier;
  */
 public class SwitchDataPlane implements ISwitchDataPlane {
 	long id;
-	private TreeMap<Short, Port> ports;
+	private TreeMap<Long, Port> ports;
 	private FlowTable flowTable;
 	private static Logger logger = Logger.getLogger("sim:");
 
-	public SwitchDataPlane(long id, TreeMap<Short, Port> ports,
+	public SwitchDataPlane(long id, TreeMap<Long, Port> ports,
 			FlowTable flowTable) {
 		super();
 		this.id = id;
@@ -49,11 +49,13 @@ public class SwitchDataPlane implements ISwitchDataPlane {
 	 * short)
 	 */
 	@Override
-	public boolean receivePkt(Packet pkt, short in_port) {
+	public boolean receivePkt(Packet pkt, long in_port) {
 		logger.info("pkt rcvd: " + pkt.toString() + " on " + in_port);
 		pkt.setIn_port(in_port);
 		PacketIdentifier pktIden = new PacketIdentifier(pkt);
+		//System.out.println("Packet Indentifier are: " + pktIden);
 		List<Action> actions = flowTable.lookup(pktIden, pkt.getSize());
+		//System.out.println("Actions are: " + actions);
 		if (actions.isEmpty()) {
 			return false;
 		} else {
@@ -69,8 +71,12 @@ public class SwitchDataPlane implements ISwitchDataPlane {
 	 * short)
 	 */
 	@Override
-	public void sendPkt(Packet pkt, short out_port) {
+	public void sendPkt(Packet pkt, long out_port) {
 		Port port = this.ports.get(out_port);
+		if(port==null){
+			System.out.println("WARNING: port " + out_port + " not found on switch " + id + ". ports: " + ports.keySet());
+			return;
+		}
 		Host host = port.getHost();
 		if (host != null) {
 			host.receivePkt(pkt);
@@ -82,7 +88,7 @@ public class SwitchDataPlane implements ISwitchDataPlane {
 	public void applyActions(Packet pkt, List<Action> actions) {
 		for (Action action : actions) {
 			if (action.getType() == ActionType.OUT_PORT) {
-				short out_port = (short) action.getValue();
+				long out_port = (long) action.getValue();
 				sendPkt(pkt, out_port);
 			}
 		}

### 2013-12-14 simulator/src/swiconsim/nwswitch/port/Port.java

@@ -9,6 +9,7 @@ public class Port {
 	Host host = null;
 	Switch sw;
 	PortStatus status;
+	public int external=0;
 
 	public Port(int id) {
 		super();
@@ -73,8 +74,12 @@ public class Port {
 
 	@Override
 	public String toString() {
-		return "Port [id=" + id + ", host=" + host.getId() + ", sw=" + sw.getId() + ", status="
+		if(host!=null){
+			return "Port [id=" + id + ", host=" + host.getId() + ", sw=" + sw.getId() + ", status="
 				+ status + "]";
+		}
+		return "Port [id=" + id + ", host=NO_HOST, sw=" + sw.getId() + ", status="
+		+ status + "]";
 	}
 	
 }

### 2013-12-14 simulator/src/swiconsim/packet/Packet.java

@@ -7,7 +7,11 @@ import swiconsim.util.IPUtil;
  * 
  *         data packet
  */
-public class Packet extends PacketIdentifier {
+public class Packet extends PacketIdentifier implements Comparable<Packet> {
+	public int nhops=0;
+	public double maxutil=0;
+	public int id=0;
+	public int last=0;
 	public int getSize() {
 		return size;
 	}
@@ -18,15 +22,36 @@ public class Packet extends PacketIdentifier {
 
 	int size;
 
-	public Packet(short in_port, int nw_src, int nw_dst, int size) {
+	public Packet(long in_port, int nw_src, int nw_dst, int size, int id) {
 		super(in_port, nw_src, nw_dst);
 		this.size = size;
+		this.nhops=0;
+		this.id=id;
+	}
+	
+	public Packet(long in_port, int nw_src, int nw_dst, int size) {
+		super(in_port, nw_src, nw_dst);
+		this.size = size;
+		this.nhops=0;
+		this.id=0;
 	}
 
 	@Override
 	public String toString() {
-		return "Packet [size=" + size + ", nw_src=" + IPUtil.toString(nw_src)
+		return "Packet [port=" + in_port +", size=" + size + ", nw_src=" + IPUtil.toString(nw_src)
 				+ ", nw_dst=" + IPUtil.toString(nw_dst) + "]";
+	}
+	
+	public boolean equals(Packet pkt){
+		if(pkt.id==id) return true;
+		return false;
+	}
+
+	@Override
+	public int compareTo(Packet packet) {
+		if(packet.id==this.id) return 0;
+		if(packet.id < this.id) return 1;
+		return -1;
 	}
 
 }

### 2013-12-14 simulator/src/swiconsim/packet/PacketIdentifier.java

@@ -8,11 +8,11 @@ import swiconsim.util.IPUtil;
  *         the fields in a pkt that have to be matched
  */
 public class PacketIdentifier {
-	short in_port;
+	long in_port;
 	int nw_src;
 	int nw_dst;
 
-	public PacketIdentifier(short in_port, int nw_src, int nw_dst) {
+	public PacketIdentifier(long in_port, int nw_src, int nw_dst) {
 		super();
 		this.in_port = in_port;
 		this.nw_src = nw_src;
@@ -25,12 +25,12 @@ public class PacketIdentifier {
 		this.nw_dst = pkt.nw_dst;
 	}
 
-	public short getIn_port() {
+	public long getIn_port() {
 		return in_port;
 	}
 
-	public void setIn_port(short in_port) {
-		this.in_port = in_port;
+	public void setIn_port(long in_port2) {
+		this.in_port = in_port2;
 	}
 
 	public int getNw_src() {

### 2013-12-14 simulator/src/swiconsim/swim/Job.java

@@ -1,0 +1,35 @@
+package swiconsim.swim;
+import swiconsim.packet.*;
+
+public class Job implements Comparable{
+	int jobid;
+	long bytes;
+	JobType type;
+	long time;
+	int src;
+	int dst;
+	public Packet packet;
+	
+	public Job(int jobid, long bytes, JobType type, long time, int src, int dst, Packet packet){
+		this.jobid = jobid;
+		this.bytes = bytes;
+		this.type = type;
+		this.time = time;
+		this.src = src;
+		this.dst = dst;
+		this.packet = packet;
+	}
+
+
+	@Override
+	public int compareTo(Object other) {
+		Job job = (Job) other;
+		if(job.time == this.time) return 0;
+		if(job.time < this.time) return 1;
+		return -1;
+	}
+	
+	public String toString(){
+		return "["+jobid +": "+bytes+": "+time+ ": " + src + "-" + dst +"]";
+	}
+}

### 2013-12-14 simulator/src/swiconsim/swim/JobType.java

@@ -1,0 +1,6 @@
+package swiconsim.swim;
+
+public enum JobType {
+	START,
+	END
+}

### 2013-12-14 simulator/src/swiconsim/swim/Simulator.java

@@ -1,0 +1,395 @@
+package swiconsim.swim;
+
+import java.io.BufferedReader;
+import java.io.BufferedWriter;
+import java.io.FileReader;
+import java.io.File;
+import java.io.FileWriter;
+import java.io.IOException;
+import java.util.Collections;
+import java.util.Iterator;
+import java.util.List;
+import java.util.Random;
+import java.util.StringTokenizer;
+import swiconsim.controller.*;
+import swiconsim.network.DataNetwork;
+import swiconsim.network.ManagementNetwork;
+import swiconsim.node.ManagementNode;
+import swiconsim.nwswitch.*;
+import swiconsim.nwswitch.port.Port;
+import swiconsim.packet.Packet;
+
+import java.util.ArrayList;
+import swiconsim.host.*;
+import java.util.HashMap;
+import swiconsim.util.*;
+
+public class Simulator {
+	
+	HashMap<ManagementNode, ArrayList<Switch>> nodes;
+	HashMap<Long, Host> hosts;
+	HashMap<Long, Short> ports;
+	Controller parent;
+	
+	BufferedWriter resultWriter;
+	BufferedWriter logWriter;
+	BufferedWriter nodeGapWriter;
+	BufferedReader nodeGapReader;
+	
+	int ndomains=10;
+	int nswitches=20;
+	int nhosts=3;
+	int nports=100;
+	int hostid=0;
+	HashMap<Long, Integer> intraNeighbors;
+	int interNeighbors=2;
+	
+	ArrayList<Job> startjobs;
+	ArrayList<Job> endjobs;
+	
+	public Simulator(String resultFile, String logFile){
+		try {
+			resultWriter = new BufferedWriter(new FileWriter(new File(resultFile)));
+			logWriter = new BufferedWriter(new FileWriter(new File(logFile)));
+		} catch (IOException e) {
+			e.printStackTrace();
+		}
+		
+	}
+	
+	void createTopology(int flag){
+		
+		if(flag==1){
+			parent = new Controller(1000);
+		}
+		
+		nodes = new HashMap<ManagementNode, ArrayList<Switch>>();
+		hosts = new HashMap<Long, Host>();
+		ports = new HashMap<Long,Short>();
+		intraNeighbors = new HashMap<Long, Integer>();
+		
+		for(int i=1; i<=ndomains; i++){
+			long cid = 10000 +i;
+			ManagementNode controller;
+			if(flag==1){
[…]

### 2013-12-14 simulator/src/swiconsim/test/SwitchTest.java

@@ -37,12 +37,12 @@ public class SwitchTest {
 		List<Action> actions = new ArrayList<Action>();
 		actions.add(new Action(ActionType.OUT_PORT, 2));
 		Flow flow = new Flow(match, actions, (short) 10);
-		Controller cont = ManagementNetwork.getInstance().getController(101);
-		cont.addFlowToSwitch(11, flow);
+		Controller cont = (Controller) ManagementNetwork.getInstance().getController(101);
+		cont.addFlowToSwitch(11, flow, null);
 		System.out.println(sw.toString());
 		
 		// Add edge 1100002 <-> 1200001
-		DataNetwork.getInstance().addEdge(11, (short)2, 12, (short)1);
+		DataNetwork.getInstance().addEdge(11, (short)2, 12, (short)1, 100);
 		
 		// A pkt that matches the installed flow
 		Packet pkt = new Packet((short) 0, IPUtil.stringToIP("1.2.3.4"), 5, 3);

### 2013-12-14 simulator/src/swiconsim/util/SwitchFlowPair.java

@@ -1,0 +1,13 @@
+package swiconsim.util;
+
+import swiconsim.flow.Flow;
+
+public class SwitchFlowPair {
+	public long sw;
+	public Flow flow;
+	
+	public SwitchFlowPair(long sw, Flow flow){
+		this.sw = sw;
+		this.flow = flow;
+	}
+}

### 2013-12-14 simulator/src/timestamp/Controller.java

@@ -128,7 +128,7 @@ public class Controller extends Thread {
 				path.remove(path.size()-1);
 			}
 		}
-		visited.remove(sw);
+		//visited.remove(sw);
 	}
 	
 	public void getBestIntraPath(int sw1, int sw2){		
@@ -166,7 +166,7 @@ public class Controller extends Thread {
 				path.remove(path.size()-1);
 			}
 		}
-		visited.remove(sw);
+		//visited.remove(sw);
 	}
 	
 	int updateUtilizations(Flow flow, ArrayList<FlowLink> path, int start, int end){
@@ -225,6 +225,10 @@ public class Controller extends Thread {
 			first=0;
 		}
 		return ret+"]";
+	}
+	
+	public static void main(String[] args){
+		
 	}
 	
 }

### 2013-12-14 paper/paper.tex
Related work section

@@
 \usepackage{url,graphicx,tabularx,array,geometry, caption, subcaption, float}
-
 %-- Commands for header
@@
 \begin{abstract}
-Network control plane, in Software Defined Networks (SDN), provides a global view of physical network to control applications (such as routing), which can be developed as a centralized application rather than a distributed one.
+Network control plane, in Software Defined Networks (SDN), provides a global view of the physical network to control applications (such as routing), which can be developed as a centralized application rather than a distributed one.
 Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple nodes called controllers, to achieve desired level of scalability, reliability and responsiveness.
 
-Traditional architecture involves multiple controllers, each operating on a subset of switches and maintaining a global view of entire network.
-We propose a different where controllers are arranged in a hierarchy and only maintain a part of network view.
+Traditional architecture involves multiple controllers, each operating on a subset of switches and maintaining a global view of the entire network.
+We propose a different architecture where controllers are arranged in a hierarchy and only view a part of the network.
 This improves responsiveness and reduces inconsistency ........(numbers).
@@
 Control plane consists of a collection of control elements, which install forwarding rules on forwarding elements as dictated by control logic or application, running on control elements.
-Some classes of control application are routing, traffic engineering (TE) and enforcing network policies.
-In traditional network, there is a tight coupling and one to one mapping between control elements and forwarding elements, with one pair on each switch.
+Some classes of control applications are routing, traffic engineering (TE), enforcing network policies, etc.
+In traditional networks, there is a tight coupling and a one to one mapping between control elements and forwarding elements, with one pair on each switch.
 Besides, the role of control elements had been trivial, which is to merely install instructions given by control application on corresponding forwarding element as forwarding rules.
-It is the control application which runs distributed protocol to synchronize the state of forwarding elements.
+It is the control application which runs distributed protocols to synchronize the state of forwarding elements.
 The design of these distributed protocols differs from application to application.
@@
 
-The emergence of Software Defined Network (SDN) has induced a separation between forwarding and control elements and has significantly improved the role of control plane.
-The control element may no longer be present on the switch but move to a separate device called controller, and uses predefined protocol such as OpenFlow [openflow-ref], to install forwarding rules.
+The emergence of Software Defined Networks (SDN) has induced a separation between forwarding and control elements and has significantly improved the role of the control plane.
+The control elements may no longer be present on the switches but move to a separate device called a controller, and use predefined protocol such as OpenFlow [openflow-ref], to install forwarding rules.
 There may be a one-to-many mapping from controllers to switches.
-At one extreme, control plane may provide control application with global network view, which can then be implemented as a centralized application, rather than a distributed one.
+At one extreme, control plane may provide a control application with the global network view, which can then be implemented as a centralized application, rather than a distributed one.
 Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple controllers, to achieve desired level of scalability, reliability and responsiveness.
 
-Conventional control plane architecture involves multiple controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application.
+Conventional control plane architectures involve multiple controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application.
 Instructions (forwarding rules) from control application are sent to target switches directly or through other controllers.
 Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT).
-Consequently, this introduces trade-off between network state inconsistency and responsiveness.
+Consequently, this introduces a trade-off between network state inconsistency and responsiveness.
 Results from [logically-centralized-ref] suggests that control state inconsistency significantly degrades performance of logically centralized control applications agnostic to the underlying state distribution.
 
-The state inconsistency problem arises from the fact that each controller try to maintain the complete global view of network.
-Besides, we believe it is difficult (from scalability point of view) for a controller (a COTS machine) to handle the entire network view.
-Motivated by these observation, we propose a hierarchical controller architecture.
+The state inconsistency problem arises from the fact that each controller tries to maintain the complete global view of the network.
+Besides, we believe that it is difficult (from scalability point of view) for a controller (a COTS machine) to handle the entire network view.
+Motivated by these observations, we propose a hierarchical controller architecture.
 We show ....................
@@
 \label{sec:hierarchical}
-We propose a control plane architecture in which controllers are arranged in the form of a tree over the physical network.
-Switches form the leaf nodes in the tree, while controllers form the internal nodes.
-An internal node acts as a switch for it's parent and acts as a controller for it's children.
-A node aggregate it's children network and converts it to a switch abstraction which is then exposed to the parent controller.
+We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network.
+Switches form the leaf nodes in the tree, while controllers form the internal nodes [TODO:tree-figure].
+An internal node acts as a switch for its parent and as a controller for its children.
+A node aggregates its underlying network and converts it to a switch abstraction which is then exposed to the parent controller.
 This switch abstraction is explained in detail in section~\ref{subsec:aggr}.
-A controller's view does not span the whole network but only the network of it's children.
-A node translates instructions (forwarding rules) from it's parent controller into forwarding rules for it's underlying network.
+A controller's view does not span the whole network but only the network of its children.
+A node translates instructions (forwarding rules) from its parent controller to the forwarding rules for its underlying network.
 Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller.
@@
 Same TE control application runs on each controller.
-When a new flow arrives, a switch sends $packet_in$ message to it's type 0 controller.
-Type 0 controller checks if the path of the flow belongs entirely to it's underlying network or not.
+When a new flow arrives, a switch sends $packet_{in}$ message to its type 0 controller.
+Type 0 controller checks if the path of the flow belongs entirely to its underlying network or not.
 If yes, it uses TE application to assign the path to the flow and no updates are sent to any controller.
-However, if the destination is outside it's domain, type 0 controller sends a $packet_in$ message to it's parent controller.
+However, if the destination is outside its domain, type 0 controller sends a $packet_{in}$ message to its parent controller.
 Parent controller uses TE control application to assign a paths of type 0 controllers.
@@
 \label{sec:related}
+Owing to the performance, reliability and scalability issues faced by single controller architectures, distributed controller architectures have been studied and implemented in various flavors.
+Most of them aim for a logically centralized global view of the network, while Some of them are discussed in this discussion.
 
+Hyperflow[Hyperflow-ref] has a logically centralized but physically distributed controller architecture.
+It localizes decision making in individual controllers by passively synchronizing the global network view.
+The control application is agnostic to the underlying architecture and takes a global decision, though the effect is local.
+The events are propagated using a publish/subscribe messaging paradigm.
+
+Onix[Onix ref] presents a platform on top of which a network control plane can be implemented as a distributed system.
+Control planes written within Onix operate on a global view of the network and use basic state distribution primitives provided by the platform.
+Thus Onix provides a general distributed state management API for control plane implementations, but leaves the decision making and trade-offs among consistency, durability and scalability to the control applications.
+This makes the design, deployment, modification and verification of control applications difficult and cost prohibitive.
+
+Kandoo[Kandoo-ref] takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the nework-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state.
+Bottom layer runs a local application for its underlying network, while the top layer subscribes to changes from the bottom layer.
+Top layer view is subscription dependent and the architecture cannot be used in a generic fasion.
+
+On a similar but different note, B4[B4-ref] uses a two level hierarchy, where the site controllers view their local network and aggregate flows to the higher level centralized controller which performs TE.
+Though the paper talks about a very particular implementation, it demonstrates in practice the benefits of the generic principle of our design.
 \section{Conclusion}

### 2013-12-15 paper/paper.tex
Section 2, first draft.

@@
 We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network.
-Switches form the leaf nodes in the tree, while controllers form the internal nodes [TODO:tree-figure].
+Switches form the leaf nodes in the tree, while controllers form the internal nodes.
 An internal node acts as a switch for its parent and as a controller for its children.
@@
 \label{subsec:aggr}
+One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions from the abstract switch to the underlying network.
+We present these separately below.
+\subsubsection{Compression}
+The primary switch abstraction comes by compressing the network into one node.
+Given an underlying network for a controller, the abstract switch exposed by it contains all the ports which are connected to switches outside the network.
+These ports form the \emph{ingress} and \emph{egress} ports for the controller.
+The switch abstraction maintains a list of abstract ports, which can be mapped to the ports of the underlying network.
+This basic abstraction is suficient for a simple control application, but we require more features for a rich and complete experience.
+A few such abstractions are presented below.
+\begin{enumerate}
+    \item \emph{Host compression}:
+    If the number of hosts is too large, and we consider each host as an external entity, individual hosts must be remembered by the root controller, which is extremely inefficient.
+Host compression can be performed in two ways, depending on the host IP distribution and the network being considered.
+    \begin{itemize}
+        \item \emph{IP compression} -
+        Similar to the forwarding rules in the actual internet, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller.
+This works well only if the assignment of switches to controllers is uniform and in accordance with prefixes.
+        \item \emph{Abstract host list} -
+        A controller can expose an abstract host or list of hosts to its parent controller.
+The parent controller places rules for the abstract hosts in the peer controllers.
+This approach would be similar to LAN, and requires address translation while entering or leaving the underlying network.
+This is more flexible, but complicated, and can be used only when IP prefix compression fails.
+    \end{itemize}
+    \item \emph{Abstract flow table}:
+        The basic flow table in a controller is comprises of the rules placed by the parent controller.
+However, for a full OpenFlow functionality, we need to support counters and queries, as in OpenFlow switches.
+These responses, again, must be an aggregation of the network.
+To this end, we must be able to efficiently represent the flows from the underlying network and obtain statistics which can be exposed to the parent.
+We propose a data structure to maintain, aggregate and update view of the flows in section~\ref{subsubsec:aggr-data}.
+\end{enumerate}
+
+\subsubsection{Translation}
+The major task of an intermediate controller is to translate messages from its parent controller to rules that can be implemented on the underlying network.
+To this end, we extend the API for the control application to take in these rules as an additional input (Note: there is no standard northbound API between a controller and a control application as of today and each controller exposes its own unique API).
+The controller uses the underlying network state and its abstract flow table to decide the rules for its children.
+It forwards the packet to its parent if and only if it is unable to handle the packet with its given knowledge.
+Typically, for a new flow,
+\begin{enumerate}
+    \item If the destination is within the underlying network, or if the abstract flow table has rules to handle the destination, the controller places rules on the underlying network to create a flow from the origin (local host or an \emph{ingress} port) of the flow to the destination (local host or an \emph{egress} port) based on its policies.
+    \item If the destination is unknown, it forwards the packet to its parent, but also holds onto a copy.
+Once the parent takes a decision and passes it as an OpenFlow rule, the controller translates it to the underlying network.
+\end{enumerate}
+
+[one-big-switch] and [pallete] propose algorithms to combine and translate end-to-end policies and routing policies to optimally place rules in switches.
+These algorithms can be used by control applications to optimally place rules, combining the abstract flow table and its local routing policies.
+
+\subsubsection{Data structures}
+\label{subsubsec:aggr-data}
+We represent the underlying network as a graph $G$ with dummy external nodes connected to each of the exposed ports.
+For each external node $e$, we create and maintain a directed flow graph $F_e$ rooted at $e$ as an overlay graph on $G$.
+$e$ here represents the \emph{ingress} port.
+End hosts and other external nodes act as sinks for these flows.
+The external nodes represent the \emph{egress} ports.
+For each pair of external nodes $e_{in}$ and $e_{out}$, we can create an aggregated flow table entry based on the flow graph $F_{e_{in}}$ for $e_{in}$.
+We propose this method to account for possible modifications in the packet headers along the paths (multiple paths are common based on routing policies) from the \emph{ingress} port to the \emph{egress} port which could prevent us from identifying same flows by just observing flows at the \emph{ingress} and \emph{egress} ports.
+Other data structures which would aid to simplify updates and aggregation are
+\begin{enumerate}
+    \item \emph{Path Information} - A list of $n$ best paths for some definition of best.
+We consider paths with least hops, but this can be varied depending on the routing policies.
+This can be used to choose the best path to route from between an \emph{ingress} and \emph{egress} port based on routing policies.
+    \item \emph{Link Information} - A list of pointers to the paths this link belongs to, and a list of flows that pass through this link.
+This can be used to obtain link utilizations to select routes with minimum maximum link utilization.
+\end{enumerate}
+One main difference between an abstract switch and an actual switch is the inequality of costs between two exposed ports.
+If parameters like hop counts are being considered, the internal routing costs must also be considered.
+This can be achieved by adding extra \emph{attributes} to pairs of external ports which can signify the internal costs which could be considered by the parent controller.
+This extra feature helps reduce the sub-optimality of this solution.
+However, it comes at the expense of needing to update costs of each pair of nodes on network changes.
+
+\subsubsection{Modifications}
+One major concern of routing is network modifications, mainly link failures.
+If the network is well connected, then an internal link failure can be handled by the local controller without affecting the parent controller.
+A parent controller only needs to concern if an exposed link is broken or if an underlying network is significantly broken, for example, if it was partitioned or the paths become extremely crowded.
+With the assumption that an underlying network capacity is greater than its exposed links, i.e.
+the internal network can handle full utilization of the exposed links, these concerns will be limited.
+
+\subsection{Hierarchical arrangement}
+The abstraction described above represents a controller at any level of the hierarchy.
+The only difference between a root controller and an internal controller is the absence of incoming OpenFLow messages in the former, while the latter receives them from its parent.
+This simplifies the control application which can be agnostic to the hierarchical level it belongs to, as long as it can support incoming OpenFlow messages.
+[TODO-tree figure].
 
@@
 \label{sec:discuss}
+Hierarchical architecture provides several benefits.
+\begin{enumerate}
+    \item Compared to a flat distributed controller architecture, this architecture will have faster flow setup times, in general.
+In a flat architecture, flow setup requires one round trip between switch and controller for each of the controller domains (say $n$) the packet travels through.
+However, in a hierarchical architecture, the packet is forwarded up to a level where the entire path is under one controller and then the rules are installed in all the controllers.
+On average, this is $O(log\;n)$ levels up and thus takes significantly less time to setup.
+Worst case, however, could be when the the two nodes are close to each other physically but belong to different controllers up till the root.
+    \item Link modifications are in general handled locally by the immediate controller and other controllers are not affected by it.
+So, there are in general no messages or state inconsistencies, as these networks do not have a global state to create inconsistencies!
+    \item As long as a control application follows the protocol, the internal routing mechanisms do not create any inconsistencies.
+In flat architecture where each controller takes a global decision but acts on a local scope, each controller must behave and take decisions exactly the same way.
+Thus, with our architecture, different controllers can run different control applications, without affecting the network.
+This has several benefits in itself.
+        \begin{enumerate}
+            \item A control application can be updated independent of the remaining network and not cause global disruptions. There might, however, be flow setup delays which are to be expected.
+            \item Different routing stategies and network policies can be implemented in different segments of the network.
+For example, one controller could route using MPLS, while another uses destination addresses.
+            \item Networks with different managements can be brought together using this architecture.
+Each network can follow its own network policies and selectively expose its \emph{ingress} and \emph{egress} ports to the parent controller.
+        \end{enumerate}
+    \item The view is simplified and thus decisions can be taken quickly, only for a small number of nodes at each level.
+\end{enumerate}
+
+This architecture, however, in not a panacea for scalable distributed controller architecture. Simplifying the network comes at the cost of losing optimality.
 

### 2013-12-15 paper/paper.tex
corrections

@@
 We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network.
-Switches form the leaf nodes in the tree, while controllers form the internal nodes.
-An internal node acts as a switch for its parent and as a controller for its children.
-A node aggregates its underlying network and converts it to a switch abstraction which is then exposed to the parent controller.
-This switch abstraction is explained in detail in section~\ref{subsec:aggr}.
-A controller's view does not span the whole network but only the network of its children.
-A node translates instructions (forwarding rules) from its parent controller to the forwarding rules for its underlying network.
-Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller.
-The translation process is explained in the following section.
+We will start by describing an important building block of the architecture, called switch abstraction.
 
@@
 \label{subsec:aggr}
-One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions from the abstract switch to the underlying network.
+A controller aggregates its underlying network and behaves as one logical switch, with huge number of ports.
+This logical switch supports control plane protocols such as OpenFlow[ref], and can be controlled by an external entity (referred to as parent controller) to install forwarding rules.
+One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions (forwarding rules) from the abstract switch to the underlying network.
 We present these separately below.
@@
 Given an underlying network for a controller, the abstract switch exposed by it contains all the ports which are connected to switches outside the network.
-These ports form the \emph{ingress} and \emph{egress} ports for the controller.
-The switch abstraction maintains a list of abstract ports, which can be mapped to the ports of the underlying network.
-This basic abstraction is suficient for a simple control application, but we require more features for a rich and complete experience.
+Such ports are referred to as \emph{external} ports.
+Conversely, ports which are connected to ports in the same network are not exposed (hidden away from abstraction) and are referred to as \emph{internal} ports.
+\emph{External} ports form the \emph{ingress} and \emph{egress} ports for the abstract switch.
+The switch abstraction maintains a list of these abstract ports, which can be (uniquely) mapped to the ports of the underlying network.
+This basic abstraction is sufficient for a simple control application, but we require more features for a rich and complete experience.
 A few such abstractions are presented below.
@@
     \item \emph{Host compression}:
-    If the number of hosts is too large, and we consider each host as an external entity, individual hosts must be remembered by the root controller, which is extremely inefficient.
+    Hosts in the underlying network are considered as external entities, which are exposed as hosts connected to the abstract switch.Ports which connect to hosts are considered to be\emph{external} ports.
+However, if the number of hosts in the underlying network is too large, exposing a list of all hosts could result in inefficient compression.
 Host compression can be performed in two ways, depending on the host IP distribution and the network being considered.
@@
         \item \emph{IP compression} -
-        Similar to the forwarding rules in the actual internet, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller.
+        Similar to the forwarding rules in an internet router, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller.
 This works well only if the assignment of switches to controllers is uniform and in accordance with prefixes.
@@
     \item \emph{Abstract flow table}:
-        The basic flow table in a controller is comprises of the rules placed by the parent controller.
+        The basic flow table in a controller is comprise of the rules placed by the parent controller.
 However, for a full OpenFlow functionality, we need to support counters and queries, as in OpenFlow switches.
@@
 \subsubsection{Translation}
-The major task of an intermediate controller is to translate messages from its parent controller to rules that can be implemented on the underlying network.
+The major task of a controller is to translate messages from its parent controller to forwarding rules that can be implemented on switches in the underlying network.
 To this end, we extend the API for the control application to take in these rules as an additional input (Note: there is no standard northbound API between a controller and a control application as of today and each controller exposes its own unique API).
@@
 
-[one-big-switch] and [pallete] propose algorithms to combine and translate end-to-end policies and routing policies to optimally place rules in switches.
+[one-big-switch ref] and [pallete ref] propose algorithms to combine and translate end-to-end policies and routing policies to optimally place rules in switches.
 These algorithms can be used by control applications to optimally place rules, combining the abstract flow table and its local routing policies.
@@
 \label{subsubsec:aggr-data}
+In this section, we present data structures, maintained by a controller to translate instructions from parent controller to children switches and to translate network updates and messages from children switches to parent controller efficiently.
+Note that this is just one of the way in which \emph{translation} can be enforced, which we have used in our implementation.
 We represent the underlying network as a graph $G$ with dummy external nodes connected to each of the exposed ports.
@@
 \subsection{Hierarchical arrangement}
+We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network.
+Switches form the leaf nodes in the tree, while controllers form the internal nodes.
+An internal node acts as a switch for its parent and as a controller for its children.
+A controller's view does not span the whole network but only the network of its children.
+A node translates instructions (forwarding rules) from its parent controller to the forwarding rules for its underlying network.
+Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller.
+
 The abstraction described above represents a controller at any level of the hierarchy.
@@
 \begin{subfigure}{.5\textwidth}
-\label{fig:path}
 \includegraphics[scale=0.3]{path}
 \caption{Application performance: path length for flows}
+\label{fig:path}
 \end{subfigure}
@@
 \begin{subfigure}{.5\textwidth}
-\label{fig:util}
 \includegraphics[scale=0.3]{util}
 \caption{Application performance: Maximum link utilization for flows}
+\label{fig:util}
 \end{subfigure}
@@
 \begin{figure}
-\label{fig:msg}
 \includegraphics[scale=0.3]{msg}
 \caption{Communication overhead per flow}
+\label{fig:msg}
 \end{figure}
@@
 While the flat architecture gives optimal results, hierarchical architecture provides sub-optimal results.
-(2) Communication overhead during flow set up, which is measured in term of number of messages passed between switches and controllers as well as within controllers (Figure ~\ref{fig:msg}).
+(2) Communication overhead during flow set up, which is measured in term of number of messages passed between switches and controllers as well as within controllers (Figure \ref{fig:msg}).
 The simulation assumes instant replication of network state between controllers.

### 2013-12-16 paper/paper.tex
Includes figure
timestamps

@@
 
+\begin{figure}
+\includegraphics[scale=0.5]{hierarchy}
+\caption{Hierarchical arrangement of controllers}
+\label{fig:hierarchy}
+\end{figure}
+
 \subsection{An example}
@@
 
-\section{Timestamps}
+\section{Timestamps: An optimization for flat controller architecture}
 \label{sec:timestamps}
-We worked hard, and achieved very little.
+In this section, we propose an optimization for flat controller architecture.
+
+\subsection{System model}
+There are multiple peer controllers, managing different parts of network and keeping a global view of entire network.
+Network states are replicated between controllers, with some replication lag i.e.
+at a given instant of time network state of two controllers may be different or inconsistent.
+Network state could include a range of characteristics of underlying network including, topology, link utilizations, routing tables on switches.
+However, for the purpose of this section, we will assume that network state only consists of link utilizations.
+Network state on a given controller consists of latest data for links that belong to it's domain, while it may contain stale data for other links.
+The topology is assumed to be static and is assumed to be known by all controllers.
+Source routing is used to route packets/flows to destination: when a new flow arrives at the ingress switch, it is sent to the parent controller, which calculates the route using control application and network state (at that point of time).
+This route is marked on the packet/flow.
+
+\subsection{Optimization}
+Let's say there are \emph{n} controller domains.
+Each controller domain, $i$ maintains a vector timestamp $T_{i}[1...n]$, where $T_{i}[j]$ contains the latest revision number of domain $j$'s network state (link utilizations), seen by domain $i$.
+Whenever links utilization is revised for links in domain $i$, the controller increments it's own revision number $T_{i}[i]$ in the timestamp.
+Timestamps are sent along with network state to other controllers for replication.
+During replication, a controller revises it's timestamp with the incoming timestamp by taking maximum for each entry between the two timestamps.
+This is a general technique for any replicated system.
+Each router is updated with the latest timestamp of it's parent controller.
+
+When a route is being set up, the controller prints it's timestamp on packet header along with the route.
+When a router in a domain $i$ receives a flow with timestamp $T_{j}$ , it checks for any index k such that $T_{j} [k] < T_{i} [k]$.
+For all such index's, it recalculate the part of the route which belongs to domain $k$ using it's own network state.
+This is because these parts of existing path were calculated on stale network state.
+
+\subsection{Evaluation}
+We used a dummy topology to show the benefits of timestamps.
+There are three controller domains 1, 2 and 3, with domain 2 connecting the other two domains.
+Flows arrive randomly between domain 1 and domain 3 and are routed through domain 2.
+Link Balancer Controller [logically - ref] runs on each controller as a control application: when a new flow arrives, a path which minimizes the maximum link utilization is assigned to the flow.
+
+\begin{figure}
+
+\begin{subfigure}{.5\textwidth}
+ \includegraphics[scale=0.4]{chart_1}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+ \includegraphics[scale=0.4]{chart_2}
+\end{subfigure}
+\caption{Comparison of Link utilizations with/without Timestamp optimization}
+\label{fig:timestamp}
+\end{figure}
+
+Results (Figure \ref{fig:timestamp}) shows that using timestamp could improve control application's performance.
+However, these are only preliminary results on a dummy topology and does not give any indication of the magnitude of improvement.
+We will evaluate timestamp optimization on a bigger topology with real world flow arrival data in the future.
 
@@
             \item A control application can be updated independent of the remaining network and not cause global disruptions. There might, however, be flow setup delays which are to be expected.
-            \item Different routing stategies and network policies can be implemented in different segments of the network.
+            \item Different routing strategies and network policies can be implemented in different segments of the network.
 For example, one controller could route using MPLS, while another uses destination addresses.

### 2013-12-16 paper/paper.tex
Conclusions and future work

--- 2013-12-15 paper/paper.tex
@@
 \label{sec:eval}
-We simulated a dummy control application on both hierarchical as well as flat controller architecture, with same network topology and flow arrival distribution.
+We simulated a dummy control application on both hierarchical and flat controller architectures, with same network topology and flow arrival distribution.
 The details of simulation, metrics and results are presented in this section.
@@
 We wrote our own simulator, which is available at [TODO REF simulator].
-We did not use an existing simulator/controller such as Mininet with Floodlight [TODO ref] for the following two reasons: (1) Existing controller lacks the support for flexible controller architecture and are not suitable for hierarchical arrangement of controllers.
-Modifying existing controllers such as Floodlight to add support for One-switch abstraction (section~\ref{subsec:aggr}), is not a good idea because it requires code/design changes in their primitive modules.
+We did not use an existing simulator/controller such as Mininet with Floodlight [TODO ref] for the following two reasons: (1) Existing controllers lack the support for flexible controller architecture and are not suitable for hierarchical arrangement of controllers.
+Modifying existing controllers such as Floodlight to add support for one-switch abstraction (section~\ref{subsec:aggr}), is not a good idea because it requires code/design changes in their primitive modules.
 Besides, we were not looking for a feature rich controller at that point of time.
 (2) Existing simulators do not provide controls to regulate replication between controllers or to measure communication overhead (number of messages exchanged ) between controllers, which we needed for evaluation of control plane architecture.
-Given the above requirements, it seemed easier to write a simulator which provides one-switch abstraction support for the controllers.
-There are three prime interfaces: (1) {\it Switch Data Plane:} It is implemented by a switch to create forwarding elements to handle packet forwarding according to forwarding rules.
-(2) {\it Switch Control Plane:} It is implemented by both controllers and switches and is used to establish connections to parent controllers to send packets which do not match any forwarding rule and to receive forwarding rules from parent controller.
-(3) {\it Controller Control Plane:} It is implemented by controllers to establish connections to children switches/controllers.
-It is used to receive packets which do not match any forwarding rule and to send forwarding rules to children switches/controllers.Simulator defines two networks: (1) {\t Data Network:} This network consists of all switches (and no controller) connected through links.
-(2) {\it Management Network:} This network consists of all controllers and switches connected through Controller - Switch Control Plane interfaces.
+Given the above requirements, writing a simulator which provides one-switch abstraction support for the controllers was the best option.
+
+There are three prime interfaces:
+\begin{enumerate}
+    \item {\it Switch Data Plane:} It is implemented by a switch to create forwarding elements to handle packet forwarding according to forwarding rules.
+    \item {\it Switch Control Plane:} It is implemented by both controllers and switches and is used to establish connections with parent controllers to send packets which do not match any forwarding rule, to receive forwarding rules and send/receive other OpenFLow-like messages from parent controller.
+    \item {\it Controller Control Plane:} It is implemented by controllers to establish connections to child switches/controllers.
+It is used to receive packets which do not match any forwarding rule and to send forwarding rules to child switches/controllers.
+\end{enumerate}
+Simulator defines two networks:
+\begin{enumerate}
+    \item {\it Data Network:} This network consists of all switches (and no controllers) connected through links.
+    \item {\it Management Network:} This network consists of all controllers and switches connected through Controller - Switch Control Plane interfaces.
 This network allows flexible controller architecture, which could be hierarchical or flat.
+\end{enumerate}
 
-To observe the effect of different architecture in real world applications, we used SWIM project [TODO ref swim], which provides synthetic representative test workloads  by sampling historical MapReduce cluster traces using [swim paper ref TODO].
-The original trace comes from historical Hadoop traces on a 600 machine cluster at Facebook and spans 6 months, and contains roughly 1 million jobs.
-The representative trace file contains ~6000 jobs.
-The trace was simulated with a cluster of 600 machines (as in the original cluster) connected through a network of 200 switches and 10 Gbps links.
-The cluster is divided into 10 control domains, with 60 hosts, 20 switches and one controller for each domain.
-Each switch in a domain is connected to 1 to 4 switches within the same domain, with degree being chosen randomly.
-Each domain contains two Top-of-the-rack switches which inter-connects domains.
-For hierarchical architecture, there are 11 controllers: 1 controller per domain and one root controllers which manages the other 10 controllers.
-While flat architecture consists of 10 peer controllers, one or each domain.
+To observe the effect of different architectures in real world applications, we used SWIM project [TODO ref swim], which provides synthetic representative test workloads  by sampling historical MapReduce cluster traces using [swim paper ref TODO].
+The original trace comes from historical Hadoop traces on a 600 machine cluster at Facebook and spans $6$ months, and contains roughly $1$ million jobs.
+The representative trace file contains $\approx 6000$ jobs.
+The trace was simulated with a cluster of $600$ machines (as in the original cluster) connected through a network of $200$ switches and $10$ Gbps links.
+The cluster is divided into $10$ control domains, with $60$ hosts, $20$ switches and one controller for each domain.
+Each switch in a domain is connected to $1$ to $4$ switches within the same domain, with degree being chosen randomly.
+Each domain contains $2$ top-of-the-rack switches which inter-connect domains.
+For hierarchical architecture, there are $11$ controllers: $1$ controller per domain and $1$ root controller which manages the other $10$ controllers.
+While flat architecture consists of $10$ peer controllers, one for each domain.
 
@@
 When a new flow arrives, control application finds the path with minimum maximum utilization and sets up forwarding rules (based on IP destination and incoming ports) along the path.
-If multiple such paths exists (which is frequently the case, due to big network), shortest path (in term of number of hops) is chosen.
+If multiple such paths exist (which is frequently the case, due to a large network), shortest path (in term of number of hops) is chosen.
 This application runs on each controller in both hierarchical and flat architecture.
-However, in the hierarchical architecture, the view of application is limited and may consists of controllers or switches as forwarding elements, while in the flat architecture, control application is presented with global view of the network.
+However, in the hierarchical architecture, the view of application is limited and may consist of controllers or switches as forwarding elements, while in the flat architecture, control application is presented with global view of the network.
 For simplicity, we have kept network topology as static i.e.
-links/nodes do not fail and/or created during the simulation.
-For each Hadoop job, a dummy MapReduce application choses directory, map and reduce nodes randomly in the network, thus, creating a number of flows which are handled by Control application running on controllers.
-The duration of flow depends on the bytes needed to transfer and assigned rates.
+links/nodes do not fail and/or are not created during the simulation.
+For each Hadoop job, a dummy MapReduce application choses a directory, map and reduce nodes randomly in the network, thus, creating a number of flows which are handled by the control application running on controllers.
+The duration of flows depends on the bytes needed to transfer and the assigned rates.
 
@@
 \label{subsec:results}
+We simulated the trace file for both flat and hierarchical architectures, with the same network topology and the same assignment of task nodes for each job. We calculated two metrics:
+\begin{enumerate}
+    \item The performance of control application which is measured as the length of the path (number of hops) assigned to flows (Figure ~\ref{fig:path}) and maximum link utilization encountered by the flow (Figure ~\ref{fig:util}).
+While the flat architecture gives optimal results, hierarchical architecture provides sub-optimal results.
+    \item Communication overhead during flow set up, which is measured in term of number of messages passed between switches and controllers as well as within controllers (Figure ~\ref{fig:msg}).
+The simulation assumes instant replication of network state between controllers.
+\end{enumerate}
+Results show that hierarchical architecture reduces the communication overhead between controllers by an order of magnitude while modestly deviating from optimal solution.
 
-\begin{figure}[t]
+\begin{figure}[h]
 \centering
@@
 
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[scale=0.3]{msg}
@@
 \end{figure}
-
-We simulated the trace file for both flat and hierarchical architecture, with same topology and same assignment of task nodes for each job.
-We calculated two metrics: (1) The performance of control application which is measured as the length path (number of hops) assigned to flows (Figure ~\ref{fig:path}) and maximum link utilization encountered by the flow (Figure ~\ref{fig:util}).
-While the flat architecture gives optimal results, hierarchical architecture provides sub-optimal results.
-(2) Communication overhead during flow set up, which is measured in term of number of messages passed between switches and controllers as well as within controllers (Figure \ref{fig:msg}).
-The simulation assumes instant replication of network state between controllers.
-Results show that hierarchical architecture reduces the communication overhead between controllers by an order of magnitude while modestly deviating from optimal solution.
 
@@
 On average, this is $O(log\;n)$ levels up and thus takes significantly less time to setup.
-Worst case, however, could be when the the two nodes are close to each other physically but belong to different controllers up till the root.
+Worst case, however, could be when the two nodes are close to each other physically but belong to different controllers up till the root.
     \item Link modifications are in general handled locally by the immediate controller and other controllers are not affected by it.
@@
 
-This architecture, however, in not a panacea for scalable distributed controller architecture. Simplifying the network comes at the cost of losing optimality.
+This architecture, however, in not a panacea for scalable distributed controller architectures as simplifying the network comes at the cost of losing optimality.
+We still believe that the reduction is performance is modest compared to the advantages it offers.
 
@@
 Owing to the performance, reliability and scalability issues faced by single controller architectures, distributed controller architectures have been studied and implemented in various flavors.
-Most of them aim for a logically centralized global view of the network, while Some of them are discussed in this discussion.
+Most of them aim for a logically centralized global view of the network, while other provide a distributed framework and leave design choices to the control application.
+Some of them are discussed in this discussion.
 
@@
 Though the paper talks about a very particular implementation, it demonstrates in practice the benefits of the generic principle of our design.
+
 \section{Conclusion}
 \label{sec:conc}
+We presented a hierarchical controller architecture where each controller exposes a switch abstraction to its parent and discussed its advantages and disadvantages.
+We implemented a simulator with a flexible architecture and simulated workloads using both hierarchical and flat distributed controller architectures.
+We demonstrated that the hierarchical architecture works very well to improve scalability and reduce communication overhead of the system, while modestly compromising on optimality.
+We believe that this architecture would be ideal to handle really large networks and hope that it will be used in the future.
+
+\section{Future work}
+In the future, we want to modify an existing OpenFlow controller like FloodLight or Open DayLight to provide support for hierarchy.
+We would also like to develop more complex control applications which can be run on these controllers.
+
+We also want to generalize and share our simulator for quick prototyping and testing of flexible architectures.
+
+We want to test our timestamps solution on real workloads and implement it as a module in existing controllers.
 

### 2013-12-16 paper/paper.tex
revised abstract and intro

@@
 \begin{abstract}
-Network control plane, in Software Defined Networks (SDN), provides a global view of the physical network to control applications (such as routing), which can be developed as a centralized application rather than a distributed one.
+Network control plane, in Software Defined Networks (SDN), provides a global view of the physical network to control applications.
 Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple nodes called controllers, to achieve desired level of scalability, reliability and responsiveness.
 
-Traditional architecture involves multiple controllers, each operating on a subset of switches and maintaining a global view of the entire network.
+Conventional control plane architecture involves multiple peer controllers, each operating on a subset of network and maintaining a global view of the entire network.
 We propose a different architecture where controllers are arranged in a hierarchy and only view a part of the network.
-This improves responsiveness and reduces inconsistency ........(numbers).
+Limiting the view of controllers significantly reduces communication overhead while modestly compromising on optimality.
+We simulated both architectures on historical Hadoop traces from Facebook and observed that hierarchical architecture reduces communication overhead by an order of magnitude (90.5\%) while deviating from optimal results by 17.4\% for routing application and 12.8\% for Traffic Engineering (TE) application.
 
-Besides, we also propose an optimization in traditional controller architecture. We show 
+Besides, we also propose an optimization in conventional flat controller architecture and shows preliminary results on a dummy topology. 
 \end{abstract}
@@
 \section{Introduction}
-Forwarding or data plane consists of a collection of forwarding elements (switches), connected to form a network, which is inherently distributed in nature.
+%Forwarding or data plane consists of a collection of forwarding elements (switches), connected to form a network, which is inherently distributed in nature.
 Control plane consists of a collection of control elements, which install forwarding rules on forwarding elements as dictated by control logic or application, running on control elements.
@@
 
-Conventional control plane architectures involve multiple controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application.
-Instructions (forwarding rules) from control application are sent to target switches directly or through other controllers.
+Conventional control plane architectures involve multiple peer controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application.
+We referred to such an architecture as \emph{flat controller architecture}.
 Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT).
-Consequently, this introduces a trade-off between network state inconsistency and responsiveness.
-Results from [logically-centralized-ref] suggests that control state inconsistency significantly degrades performance of logically centralized control applications agnostic to the underlying state distribution.
+Consequently, such an architecture results in huge communication overhead between controllers, which may hinders the scalability of the system.
+Besides, network state inconsistency is quite common in flat architecture which degrades the application performance significantly [logically ref TODO].
 
-The state inconsistency problem arises from the fact that each controller tries to maintain the complete global view of the network.
-Besides, we believe that it is difficult (from scalability point of view) for a controller (a COTS machine) to handle the entire network view.
-Motivated by these observations, we propose a hierarchical controller architecture.
-We show ....................
+The state inconsistency and huge communication overhead problem arises from the fact that each controller tries to maintain the complete global view of the network.
+Motivated by this observation, we propose a hierarchical controller architecture, in which controllers are arranged in a tree structure and exposes a switch abstraction to it's parent.
+We show that hierarchical architecture significantly reduces the communication overhead between controllers while modestly compromising on optimality.
 
@@
 Section~\ref{sec:hierarchical} presents Hierarchical Controller Architecture.
-We present the experimental results in section~\ref{sec:eval}.
-In section~\ref{sec:timestamps}, we present an optimization to conventional controller architecture to help improve the performance of application even if states between controllers is inconsistent.
-Related work is presented in section~\ref{sec:related}.
-We discuss ............Our new and exciting results are described in Section~\ref{sec:eval}.
+We discuss implementation details and present experimental results in section~\ref{sec:eval}.
+In section~\ref{sec:timestamps}, we present an optimization to conventional flat controller architecture.
+We discuss pros and cons of proposed architecture in section \ref{sec:discuss}.
+Related work is discussed in section~\ref{sec:related}.
+We conclude the paper in section \ref{sec:conc}, followed by future work in section \ref{sec:future}.
 
@@
 \centering
-\begin{tabular}{| L{5cm} | L{3cm} | L{3cm} | L{5cm} |}
+\begin{tabular}{| L{4cm} | L{2cm} | L{3cm} | L{4cm} |}
 \hline
@@
 \end{tabular}
+\caption{Comparison between performance of Flat and Hierarchical architecture.}
 \end{table*}%
@@
 \section{Future work}
+\label{sec:future}
 In the future, we want to modify an existing OpenFlow controller like FloodLight or Open DayLight to provide support for hierarchy.
@@
 
-We want to test our timestamps solution on real workloads and implement it as a module in existing controllers.
+We want to test our timestamp optimization on real workloads and implement it as a module in existing controllers.
 

### 2013-12-16 paper/paper.tex
references
refs

@@
 A controller aggregates its underlying network and behaves as one logical switch, with huge number of ports.
-This logical switch supports control plane protocols such as OpenFlow[ref], and can be controlled by an external entity (referred to as parent controller) to install forwarding rules.
+This logical switch supports control plane protocols such as OpenFlow\cite{openflow,of-spec}, and can be controlled by an external entity (referred to as parent controller) to install forwarding rules.
 One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions (forwarding rules) from the abstract switch to the underlying network.
@@
 
-[one-big-switch ref] and [pallete ref] propose algorithms to combine and translate end-to-end policies and routing policies to optimally place rules in switches.
+\cite{one-big-switch} proposes algorithms to combine and translate end-to-end policies and routing policies to optimally place rules in switches.
 These algorithms can be used by control applications to optimally place rules, combining the abstract flow table and its local routing policies.
@@
 \subsection{Hierarchical arrangement}
-We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network.
+We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network as shown in figure ~\ref{fig:hierarchy}.
 Switches form the leaf nodes in the tree, while controllers form the internal nodes.
@@
 This simplifies the control application which can be agnostic to the hierarchical level it belongs to, as long as it can support incoming OpenFlow messages.
-[TODO-tree figure].
 
@@
 
-Hyperflow[Hyperflow-ref] has a logically centralized but physically distributed controller architecture.
+Hyperflow\cite{hyperflow} has a logically centralized but physically distributed controller architecture.
 It localizes decision making in individual controllers by passively synchronizing the global network view.
@@
 
-Onix[Onix ref] presents a platform on top of which a network control plane can be implemented as a distributed system.
+Onix\cite{onix} presents a platform on top of which a network control plane can be implemented as a distributed system.
 Control planes written within Onix operate on a global view of the network and use basic state distribution primitives provided by the platform.
@@
 
-Kandoo[Kandoo-ref] takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the nework-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state.
+Kandoo\cite{kandoo} takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the nework-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state.
 Bottom layer runs a local application for its underlying network, while the top layer subscribes to changes from the bottom layer.
@@
 
-On a similar but different note, B4[B4-ref] uses a two level hierarchy, where the site controllers view their local network and aggregate flows to the higher level centralized controller which performs TE.
+On a similar but different note, B4\cite{b4} uses a two level hierarchy, where the site controllers view their local network and aggregate flows to the higher level centralized controller which performs TE.
 Though the paper talks about a very particular implementation, it demonstrates in practice the benefits of the generic principle of our design.
@@
 \label{sec:ack}
+\nocite{*}
 
 \bibliographystyle{abbrv}
-\bibliography{main}
+\bibliography{paper}
 

### 2013-12-16 paper/paper.tex
references

@@
 The emergence of Software Defined Networks (SDN) has induced a separation between forwarding and control elements and has significantly improved the role of the control plane.
-The control elements may no longer be present on the switches but move to a separate device called a controller, and use predefined protocol such as OpenFlow [openflow-ref], to install forwarding rules.
+The control elements may no longer be present on the switches but move to a separate device called a controller, and use predefined protocol such as OpenFlow \cite{openflow}, to install forwarding rules.
 There may be a one-to-many mapping from controllers to switches.
@@
 Consequently, such an architecture results in huge communication overhead between controllers, which may hinders the scalability of the system.
-Besides, network state inconsistency is quite common in flat architecture which degrades the application performance significantly [logically ref TODO].
+Besides, network state inconsistency is quite common in flat architecture which degrades the application performance significantly \cite{logically-centralized}.
 
@@
 \includegraphics[scale=0.5]{hierarchy}
-\caption{Hierarchical arrangement of controllers}
+\caption{A 2-layer Hierarchical arrangement of controllers: Root controller manages the layer of controllers, which in-turn manages a subset of switches.}
 \label{fig:hierarchy}
@@
 \label{sec:implementation}
-We wrote our own simulator, which is available at [TODO REF simulator].
-We did not use an existing simulator/controller such as Mininet with Floodlight [TODO ref] for the following two reasons: (1) Existing controllers lack the support for flexible controller architecture and are not suitable for hierarchical arrangement of controllers.
+We wrote our own simulator, which is available at \cite{simulator}.
+We did not use an existing simulator/controller such as Mininet\cite{mininet} with Floodlight\cite{floodlight} for the following two reasons: (1) Existing controllers lack the support for flexible controller architecture and are not suitable for hierarchical arrangement of controllers.
 Modifying existing controllers such as Floodlight to add support for one-switch abstraction (section~\ref{subsec:aggr}), is not a good idea because it requires code/design changes in their primitive modules.
@@
 
-To observe the effect of different architectures in real world applications, we used SWIM project [TODO ref swim], which provides synthetic representative test workloads  by sampling historical MapReduce cluster traces using [swim paper ref TODO].
+To observe the effect of different architectures in real world applications, we used SWIM project \cite{swim}, which provides synthetic representative test workloads  by sampling historical MapReduce cluster traces.
 The original trace comes from historical Hadoop traces on a 600 machine cluster at Facebook and spans $6$ months, and contains roughly $1$ million jobs.
@@
 While the flat architecture gives optimal results, hierarchical architecture provides sub-optimal results.
+However, note that simulation provides instantaneous replication and thus eliminates network inconsistency between controller, which would  have otherwise, degraded the performance of flat architecture.
     \item Communication overhead during flow set up, which is measured in term of number of messages passed between switches and controllers as well as within controllers (Figure ~\ref{fig:msg}).
@@
 Flows arrive randomly between domain 1 and domain 3 and are routed through domain 2.
-Link Balancer Controller [logically - ref] runs on each controller as a control application: when a new flow arrives, a path which minimizes the maximum link utilization is assigned to the flow.
+Link Balancer Controller \cite{logically-centralized} runs on each controller as a control application: when a new flow arrives, a path which minimizes the maximum link utilization is assigned to the flow.
 

### 2013-12-16 paper/paper.tex
corrected typos...

@@
         \item \emph{IP compression} -
-        Similar to the forwarding rules in an internet router, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller.
+        Similar to the forwarding rules in an Internet router, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller.
 This works well only if the assignment of switches to controllers is uniform and in accordance with prefixes.
@@
 The abstraction described above represents a controller at any level of the hierarchy.
-The only difference between a root controller and an internal controller is the absence of incoming OpenFLow messages in the former, while the latter receives them from its parent.
+The only difference between a root controller and an internal controller is the absence of incoming OpenFlow messages in the former, while the latter receives them from its parent.
 This simplifies the control application which can be agnostic to the hierarchical level it belongs to, as long as it can support incoming OpenFlow messages.
@@
     \item {\it Switch Data Plane:} It is implemented by a switch to create forwarding elements to handle packet forwarding according to forwarding rules.
-    \item {\it Switch Control Plane:} It is implemented by both controllers and switches and is used to establish connections with parent controllers to send packets which do not match any forwarding rule, to receive forwarding rules and send/receive other OpenFLow-like messages from parent controller.
+    \item {\it Switch Control Plane:} It is implemented by both controllers and switches and is used to establish connections with parent controllers to send packets which do not match any forwarding rule, to receive forwarding rules and send/receive other OpenFlow-like messages from parent controller.
     \item {\it Controller Control Plane:} It is implemented by controllers to establish connections to child switches/controllers.
@@
 
-Kandoo\cite{kandoo} takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the nework-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state.
+Kandoo\cite{kandoo} takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the network-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state.
 Bottom layer runs a local application for its underlying network, while the top layer subscribes to changes from the bottom layer.
-Top layer view is subscription dependent and the architecture cannot be used in a generic fasion.
+Top layer view is subscription dependent and the architecture cannot be used in a generic fashion.
 

### 2013-12-16 paper/paper.tex
minor changes

--- 2013-12-16 paper/paper.tex
@@
 Some classes of control applications are routing, traffic engineering (TE), enforcing network policies, etc.
-In traditional networks, there is a tight coupling and a one to one mapping between control elements and forwarding elements, with one pair on each switch.
+In traditional networks, there is a tight coupling and a one-to-one mapping between control elements and forwarding elements, with one pair on each switch.
 Besides, the role of control elements had been trivial, which is to merely install instructions given by control application on corresponding forwarding element as forwarding rules.
@@
 Conventional control plane architectures involve multiple peer controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application.
-We referred to such an architecture as \emph{flat controller architecture}.
+We refer to such an architecture as \emph{flat controller architecture}.
 Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT).
@@
 The state inconsistency and huge communication overhead problem arises from the fact that each controller tries to maintain the complete global view of the network.
-Motivated by this observation, we propose a hierarchical controller architecture, in which controllers are arranged in a tree structure and exposes a switch abstraction to it's parent.
+Motivated by this observation, we propose a hierarchical controller architecture, in which controllers are arranged in a tree structure and exposes a switch abstraction to its parent.
 We show that hierarchical architecture significantly reduces the communication overhead between controllers while modestly compromising on optimality.

### 2013-12-16 paper/paper.tex
correction

--- 2013-12-16 paper/paper.tex
@@
 Same TE control application runs on each controller.
-When a new flow arrives, a switch sends $packet_{in}$ message to its type 0 controller.
+When a new flow arrives, a switch sends $packet\_in$ message to its type 0 controller.
 Type 0 controller checks if the path of the flow belongs entirely to its underlying network or not.
 If yes, it uses TE application to assign the path to the flow and no updates are sent to any controller.
-However, if the destination is outside its domain, type 0 controller sends a $packet_{in}$ message to its parent controller.
+However, if the destination is outside its domain, type 0 controller sends a $packet\_in$ message to its parent controller.
 Parent controller uses TE control application to assign a paths of type 0 controllers.
@@
 If the path length is $l$, there will be a maximum of $l+1$ forwarding rule messages between controllers and no link utilization updates.
+However, the path assigned might not be the one with minimum maximum utilization.
 This significant reduction in communication overhead comes at the cost of sub-optimality in the application performance.

### 2013-12-16 paper/paper.tex
minor changes

--- 2013-12-16 paper/paper.tex
@@
     \item \emph{Host compression}:
-    Hosts in the underlying network are considered as external entities, which are exposed as hosts connected to the abstract switch.Ports which connect to hosts are considered to be\emph{external} ports.
+    Hosts in the underlying network are considered as external entities, which are exposed as hosts connected to the abstract switch.
+Ports which connect to hosts are considered to be \emph{external} ports.
 However, if the number of hosts in the underlying network is too large, exposing a list of all hosts could result in inefficient compression.
@@
     \item \emph{Abstract flow table}:
-        The basic flow table in a controller is comprise of the rules placed by the parent controller.
+        The basic flow table in a controller comprises of the rules placed by the parent controller.
 However, for a full OpenFlow functionality, we need to support counters and queries, as in OpenFlow switches.
@@
 \label{subsubsec:aggr-data}
-In this section, we present data structures, maintained by a controller to translate instructions from parent controller to children switches and to translate network updates and messages from children switches to parent controller efficiently.
-Note that this is just one of the way in which \emph{translation} can be enforced, which we have used in our implementation.
+In this section, we present data structures, maintained by a controller to translate instructions from parent controller to child switches and to translate network updates and messages from child switches to parent controller efficiently.
+Note that this is just one of the ways in which \emph{translation} can be enforced, which we have used in our implementation.
 We represent the underlying network as a graph $G$ with dummy external nodes connected to each of the exposed ports.
@@
 \subsection{System model}
-There are multiple peer controllers, managing different parts of network and keeping a global view of entire network.
+There are multiple peer controllers, managing different parts of the network and keeping a global view of the entire network.
 Network states are replicated between controllers, with some replication lag i.e.
 at a given instant of time network state of two controllers may be different or inconsistent.
-Network state could include a range of characteristics of underlying network including, topology, link utilizations, routing tables on switches.
+Network state could include a range of characteristics of underlying network including, topology, link utilizations, routing tables on switches, etc.
 However, for the purpose of this section, we will assume that network state only consists of link utilizations.

### 2013-12-16 paper/paper.tex
minor changes

@@
 Each controller domain, $i$ maintains a vector timestamp $T_{i}[1...n]$, where $T_{i}[j]$ contains the latest revision number of domain $j$'s network state (link utilizations), seen by domain $i$.
-Whenever links utilization is revised for links in domain $i$, the controller increments it's own revision number $T_{i}[i]$ in the timestamp.
-Timestamps are sent along with network state to other controllers for replication.
-During replication, a controller revises it's timestamp with the incoming timestamp by taking maximum for each entry between the two timestamps.
+Whenever links utilization is revised for links in domain $i$, the controller increments its own revision number $T_{i}[i]$ in the timestamp.
+Timestamps are sent along with the network state to other controllers for replication.
+During replication, a controller revises its timestamp with the incoming timestamp by taking maximum for each entry between the two timestamps.
 This is a general technique for any replicated system.
-Each router is updated with the latest timestamp of it's parent controller.
+Each router is updated with the latest timestamp of its parent controller.
 
-When a route is being set up, the controller prints it's timestamp on packet header along with the route.
-When a router in a domain $i$ receives a flow with timestamp $T_{j}$ , it checks for any index k such that $T_{j} [k] < T_{i} [k]$.
-For all such index's, it recalculate the part of the route which belongs to domain $k$ using it's own network state.
-This is because these parts of existing path were calculated on stale network state.
+When a route is being set up, the controller prints its timestamp on packet header along with the route.
+When a router in a domain $i$ receives a flow with timestamp $T_{j}$ , it checks for any index $k$ such that $T_{j} [k] < T_{i} [k]$.
+For all such indices, it recalculates the part of the route which belongs to domain $k$ using its own network state.
+This is because these parts of the existing path were calculated on a stale network state.
 

### 2013-12-16 paper/paper.tex
minor correction
ack

@@
 However, these are only preliminary results on a dummy topology and does not give any indication of the magnitude of improvement.
-We will evaluate timestamp optimization on a bigger topology with real world flow arrival data in the future.
+Going ahead, we plan to evaluate timestamp optimization on a bigger topology with real world flow arrival data.
 
@@
 \label{sec:ack}
+We would like to thank Prof. Brighten Godfrey and Prof. Matthew Caesar for their useful suggestions and support. 
 \nocite{*}

### 2013-12-16 paper/paper.tex
authors
mini minor change

@@
 \author{
-        Author 1 \\
+        Gourav Khaneja \\
+	UIUC \\
+	khaneja2@illinois.edu \\
             \and
-        Author 2\\
-        %UIUC\\
+        Sweta Seethamraju\\
+        UIUC\\
+	seetham2@illinois.edu \\
 			\and
-		Author 3\\
+	Praveen Kumar\\
+	IBM \\
+	praveek7@in.ibm.com \\
 }
-\date{\today}
 
+\date{}
 \documentclass[10pt, twocolumn]{article}
@@
 Each controller domain, $i$ maintains a vector timestamp $T_{i}[1...n]$, where $T_{i}[j]$ contains the latest revision number of domain $j$'s network state (link utilizations), seen by domain $i$.
-Whenever links utilization is revised for links in domain $i$, the controller increments its own revision number $T_{i}[i]$ in the timestamp.
+Whenever link utilizations are revised for links in domain $i$, the controller increments its own revision number $T_{i}[i]$ in the timestamp.
 Timestamps are sent along with the network state to other controllers for replication.

### 2013-12-16 poster/poster.tex
implementation and results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% a0poster Portrait Poster
% LaTeX Template
% Version 1.0 (22/06/13)
%
% The a0poster class was created by:
% Gerlinde Kettl and Matthias Weiser (tex@kettl.de)
% 
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0,portrait]{a0poster}

\usepackage{multicol} % This is so we can have multiple columns of text side-by-side
\columnsep=100pt % This is the amount of white space between the columns in the poster
\columnseprule=3pt % This is the thickness of the black line between the columns in the poster

\usepackage[svgnames]{xcolor} % Specify colors by their 'svgnames', for a full list of all colors available see here: http://www.latextemplates.com/svgnames-colors

\usepackage{times} % Use the times font
%\usepackage{palatino} % Uncomment to use the Palatino font

\usepackage{graphicx} % Required for including images
\graphicspath{{figures/}} % Location of the graphics files
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{amsfonts, amsmath, amsthm, amssymb} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures
%\usepackage{subfig}
\usepackage{float}
\usepackage[it]{subfigure}

\begin{document}

%----------------------------------------------------------------------------------------
%	POSTER HEADER 
%----------------------------------------------------------------------------------------

% The header is divided into two boxes:
% The first is 75% wide and houses the title, subtitle, names, university/organization and contact information
% The second is 25% wide and houses a logo for your university/organization or a photo of you
% The widths of these boxes can be easily edited to accommodate your content as you see fit

\begin{minipage}[b]{0.75\linewidth}
\veryHuge \color{NavyBlue} \textbf{Hierarchical Controller Architecture} \color{Black}\\ % Title
\Huge\color{NavyBlue}\text{in Software Defined Networks}\\[1cm] % Subtitle
\huge \textbf{Author1 \& Author2}\\[0.5cm] % Author(s)
%\huge University and Department Name\\[0.4cm] % University/organization
%\Large \texttt{john@LaTeXTemplates.com} --- 1 (000) 111 1111\\
\end{minipage}
%
\begin{minipage}[b]{0.25\linewidth}
\includegraphics[width=20cm]{logo.png}\\
\end{minipage}

\vspace{1cm} % A bit of extra whitespace between the header and poster content

%----------------------------------------------------------------------------------------

\begin{multicols}{2} % This is how many columns your poster will be broken into, a portrait poster is generally split into 2 columns

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\color{Navy} % Navy color for the abstract

\begin{abstract}

Sed fringilla tempus hendrerit. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Etiam ut elit sit amet metus lobortis consequat sit amet in libero. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Phasellus vel sem magna. Nunc at convallis urna. isus ante. Pellentesque condimentum dui. Etiam sagittis purus non tellus tempor volutpat. Donec et dui non massa tristique adipiscing. Quisque vestibulum eros eu. Phasellus imperdiet, tortor vitae congue bibendum, felis enim sagittis lorem, et volutpat ante orci sagittis mi. Morbi rutrum laoreet semper. Morbi accumsan enim nec tortor consectetur non commodo nisi sollicitudin. Proin sollicitudin. Pellentesque eget orci eros. Fusce ultricies, tellus et pellentesque fringilla, ante massa luctus libero, quis tristique purus urna nec nibh.

\end{abstract}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\color{SaddleBrown} % SaddleBrown color for the introduction

\section*{Introduction}

Aliquam non lacus dolor, \textit{a aliquam quam} \cite{Smith:2012qr}. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Nulla in nibh mauris. Donec vel ligula nisi, a lacinia arcu. Sed mi dui, malesuada vel consectetur et, egestas porta nisi. Sed eleifend pharetra dolor, et dapibus est vulputate eu. \textbf{Integer faucibus elementum felis vitae fringilla.} In hac habitasse platea dictumst. Duis tristique rutrum nisl, nec vulputate elit porta ut. Donec sodales sollicitudin turpis sed convallis. Etiam mauris ligula, blandit adipiscing condimentum eu, dapibus pellentesque risus.

\textit{Aliquam auctor}, metus id ultrices porta, risus enim cursus sapien, quis iaculis sapien tortor sed odio. Mauris ante orci, euismod vitae tincidunt eu, porta ut neque. Aenean sapien est, viverra vel lacinia nec, venenatis eu nulla. Maecenas ut nunc nibh, et tempus libero. Aenean vitae risus ante. Pellentesque condimentum dui. Etiam sagittis purus non tellus tempor volutpat. Donec et dui non massa tristique adipiscing.

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\color{DarkSlateGray} % DarkSlateGray color for the rest of the content

\section*{Main Objectives}

\begin{enumerate}
\item Lorem ipsum dolor sit amet, consectetur.
\item Nullam at mi nisl. Vestibulum est purus, ultricies cursus volutpat sit amet, vestibulum eu.
\item Praesent tortor libero, vulputate quis elementum a, iaculis.
\item Phasellus a quam mauris, non varius mauris. Fusce tristique, enim tempor varius porta, elit purus commodo velit, pretium mattis ligula nisl nec ante.
\item Ut adipiscing accumsan sapien, sit amet pretium.
\item Estibulum est purus, ultricies cursus volutpat
\item Nullam at mi nisl. Vestibulum est purus, ultricies cursus volutpat sit amet, vestibulum eu.
\item Praesent tortor libero, vulputate quis elementum a, iaculis.
\end{enumerate}

%----------------------------------------------------------------------------------------
%	MATERIALS AND METHODS
%----------------------------------------------------------------------------------------

\section*{Materials and Methods}

Fusce magna risus, molestie ut porttitor in, consectetur sed mi. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque consectetur blandit pellentesque. Sed odio justo, viverra nec porttitor vel, lacinia a nunc. Suspendisse pulvinar euismod arcu, sit amet accumsan enim fermentum quis. In id mauris ut dui feugiat egestas. Vestibulum ac turpis lacinia nisl commodo sagittis eget sit amet sapien.

%------------------------------------------------

\subsection*{Mathematical Section}

Nulla vel nisl sed mauris auctor mollis non sed.

\begin{equation}
E = mc^{2}
\label{eqn:Einstein}
\end{equation}

Curabitur mi sem, pulvinar quis aliquam rutrum. (1) edf (2)
, $\Omega=[-1,1]^3$, maecenas leo est, ornare at. $z=-1$ edf $z=1$ sed interdum felis dapibus sem. $x$ set $y$ ytruem. 
Turpis $j$ amet accumsan enim $y$-lacina; 
ref $k$-viverra nec porttitor $x$-lacina.

Vestibulum ac diam a odio tempus congue. Vivamus id enim nisi:

\begin{eqnarray}
\cos\bar{\phi}_k Q_{j,k+1,t} + Q_{j,k+1,x}+\frac{\sin^2\bar{\phi}_k}{T\cos\bar{\phi}_k} Q_{j,k+1} &=&\nonumber\\ 
-\cos\phi_k Q_{j,k,t} + Q_{j,k,x}-\frac{\sin^2\phi_k}{T\cos\phi_k} Q_{j,k}\label{edgek}
\end{eqnarray}
and
\begin{eqnarray}
\cos\bar{\phi}_j Q_{j+1,k,t} + Q_{j+1,k,y}+\frac{\sin^2\bar{\phi}_j}{T\cos\bar{\phi}_j} Q_{j+1,k}&=&\nonumber \\
-\cos\phi_j Q_{j,k,t} + Q_{j,k,y}-\frac{\sin^2\phi_j}{T\cos\phi_j} Q_{j,k}.\label{edgej}
\end{eqnarray}

Nulla sed arcu arcu. Duis et ante gravida orci venenatis tincidunt. Fusce vitae lacinia metus. Pellentesque habitant morbi. $\mathbf{A}\underline{\xi}=\underline{\beta}$ Vim $\underline{\xi}$ enum nidi $3(P+2)^{2}$ lacina. Id feugain $\mathbf{A}$ nun quis; magno.

%Gourav - I'm starting from here ownwards: 
%----------------------------------------------------------------------------------------
%	RESULTS 
%----------------------------------------------------------------------------------------

\section*{Experimental set up}

\begin{itemize}
%\item We simulated a dummy control application on both hierarchical and flat controller architectures, with same network topology and flow arrival distribution.
\item Flow data (6000 jobs) was generated by sampling historical Hadoop traces on a 600 machine cluster at Facebook (1 million jobs). A dummy MapReduce application choses directory, map and reduce nodes randonly for each job.
\item Topology: 600 hosts connected through a network of 200 switches and 10 Gbps links. The cluster is divided into 10 equally-sized control domains.
\item TE-cum-routing coutrol application: When a new flow arrives, it finds the path with minimum maximum utilization. If multiple such paths exist (which is frequently the case, due to a large network), shortest path (in term of number of hops) is chosen.
\end{itemize}

\section*{Metrices}
\begin{itemize}
\item Path length (number of hops) of the routes assigned to flows.
\item Maximum link utilization of the path assigned to flows.
\item Communication overhead (number of messages exchanged between controllers) for seting up routes for flows.   
\end{itemize}

%\begin{center}\vspace{1cm}
%\includegraphics[width=0.8\linewidth]{msg}
%\captionof{figure}{\color{Green} Per flow communication overhead}
%\label{fig:msg}
%\end{center}\vspace{1cm}
\begin{figure}[H]
\centering
\includegraphics[scale=1]{msg}
\caption{Per flow communication overhead}
\label{fig:msg}
\end{figure}

%\begin{center}\vspace{1cm}
%\includegraphics[width=0.5\linewidth]{path}
%\captionof{figure}{\color{Green} Path length (number of hops) assigned to flows}
%\end{center}\vspace{1cm}

%\begin{center}\vspace{1cm}
%\includegraphics[width=0.5\linewidth]{util}
%\captionof{figure}{\color{Green} Maximum link utilization encountered by flows}
%\end{center}\vspace{1cm}

\begin{figure}[H]
  \begin{center}  
    \subfigure[Path length]{\label{fig:path}\includegraphics[scale=0.75]{path.png}} 
    \subfigure[Maximum link utilization]{\label{fig:util}\includegraphics[scale=0.75]{util.png}} 
  \end{center}
\caption{Application performance}
\end{figure}

\section*{Results}
Hierarchical architecture reduces communication overhead (Figure \ref{fig:msg}) by an order of magnitude while comprimising modestly on optimality: shortest path length (17.4\%) (Figure \ref{fig:path}) and minimum maximum link utilization (12.8\%) (Figure \ref{fig:util}).

%----------------------------------------------------------------------------------------
%	CONCLUSIONS
%----------------------------------------------------------------------------------------

%\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out

%\section*{Conclusions}

\color{DarkSlateGray} % Set the color back to DarkSlateGray for the rest of the content

%----------------------------------------------------------------------------------------
%	FORTHCOMING RESEARCH
%----------------------------------------------------------------------------------------

%\section*{Forthcoming Research}

 %----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\nocite{*} % Print all references regardless of whether they were cited in the poster or not
\bibliographystyle{plain} % Plain referencing style
\bibliography{sample} % Use the example bibliography file sample.bib

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\section*{Source code}
https://github.com/MugiwaraLuffy/ACN/

%----------------------------------------------------------------------------------------

\end{multicols}
\end{document}

### 2013-12-16 poster/poster.tex
small correction

@@
 %\item We simulated a dummy control application on both hierarchical and flat controller architectures, with same network topology and flow arrival distribution.
-\item Flow data (6000 jobs) was generated by sampling historical Hadoop traces on a 600 machine cluster at Facebook (1 million jobs).
-A dummy MapReduce application choses directory, map and reduce nodes randonly for each job.
+\item Flow data (6000 MapReduce jobs) was generated by sampling historical Hadoop traces on a 600 machine cluster at Facebook (1 million jobs).
+A dummy MapReduce application choses directory, map and reduce nodes randomly for each job.
 \item Topology: 600 hosts connected through a network of 200 switches and 10 Gbps links. The cluster is divided into 10 equally-sized control domains.

### 2013-12-17 poster/poster.tex
small correction
first half initial draft

@@
 
-\color{Navy} % Navy color for the abstract
-
-\begin{abstract}
-
-Sed fringilla tempus hendrerit.
+%\color{Navy} % Navy color for the abstract
+%
+%\begin{abstract}
+%
+%Sed fringilla tempus hendrerit.
 Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Etiam ut elit sit amet metus lobortis consequat sit amet in libero.
@@
 Fusce ultricies, tellus et pellentesque fringilla, ante massa luctus libero, quis tristique purus urna nec nibh.
-
-\end{abstract}
+%
+%\end{abstract}
 
@@
 
-\color{SaddleBrown} % SaddleBrown color for the introduction
-
-\section*{Introduction}
-
-Aliquam non lacus dolor, \textit{a aliquam quam} \cite{Smith:2012qr}.
+%\color{SaddleBrown} % SaddleBrown color for the introduction
+%
+%\section*{Introduction}
+%
+%Aliquam non lacus dolor, \textit{a aliquam quam} \cite{Smith:2012qr}.
 Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
@@
 Etiam mauris ligula, blandit adipiscing condimentum eu, dapibus pellentesque risus.
-
-\textit{Aliquam auctor}, metus id ultrices porta, risus enim cursus sapien, quis iaculis sapien tortor sed odio.
+%
+%\textit{Aliquam auctor}, metus id ultrices porta, risus enim cursus sapien, quis iaculis sapien tortor sed odio.
 Mauris ante orci, euismod vitae tincidunt eu, porta ut neque.
@@
 Donec et dui non massa tristique adipiscing.
-
+%
 %----------------------------------------------------------------------------------------
@@
 
-\section*{Main Objectives}
+\section*{Problem}
 
-\begin{enumerate}
-\item Lorem ipsum dolor sit amet, consectetur.
-\item Nullam at mi nisl. Vestibulum est purus, ultricies cursus volutpat sit amet, vestibulum eu.
-\item Praesent tortor libero, vulputate quis elementum a, iaculis.
-\item Phasellus a quam mauris, non varius mauris. Fusce tristique, enim tempor varius porta, elit purus commodo velit, pretium mattis ligula nisl nec ante.
-\item Ut adipiscing accumsan sapien, sit amet pretium.
-\item Estibulum est purus, ultricies cursus volutpat
-\item Nullam at mi nisl. Vestibulum est purus, ultricies cursus volutpat sit amet, vestibulum eu.
-\item Praesent tortor libero, vulputate quis elementum a, iaculis.
-\end{enumerate}
+\begin{tabular} {l | l}
+    \textbf{Single Controller Architecture} & \textbf{Distributed controller Architecture} \\
+    \hline
+    Responsiveness & Communication overhead \\
+       Scalability & Scalability of global network view \\
+       Reliability & Network state inconsistencies \\
+                   & Management barriers across networks
+\end{tabular}
 
-%----------------------------------------------------------------------------------------
-%	MATERIALS AND METHODS
-%----------------------------------------------------------------------------------------
+\textbf{Trade-offs}
+\begin{itemize}
+    \item Consistency vs. Responsiveness
+    \item Optimality vs. I dont know what
+    \item Application design complexity (Architecture aware vs. agnostic)
+\end{itemize}
 
-\section*{Materials and Methods}
+\textbf{Our Goals}
+\begin{itemize}
+    \item Make control applications agnostic to architecture
+    \item Reduce communication overhead
+    \item Minimize network state inconsistencies
+    \item Enable smooth transition across management barriers
+    \item Handle changes locally
+\end{itemize}
 
-Fusce magna risus, molestie ut porttitor in, consectetur sed mi.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque consectetur blandit pellentesque.
-Sed odio justo, viverra nec porttitor vel, lacinia a nunc.
-Suspendisse pulvinar euismod arcu, sit amet accumsan enim fermentum quis.
-In id mauris ut dui feugiat egestas.
-Vestibulum ac turpis lacinia nisl commodo sagittis eget sit amet sapien.
+\textbf{Price we pay}
+\begin{itemize}
+    \item Solution not always optimal
+\end{itemize}
+
+\section*{Our solution}
+\begin{itemize}
+    \item Hierarchical arrangement of controllers
+    \item One switch abstraction
+    \item Translate messages across hierarchical layers
+\end{itemize}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=2]{hierarchy}
+\caption{Hierarchical arrangement of controllers}
+\label{fig:hierarchy}
+\end{figure}
 
 %------------------------------------------------
-
-\subsection*{Mathematical Section}
-
-Nulla vel nisl sed mauris auctor mollis non sed.
-
-\begin{equation}
-E = mc^{2}
-\label{eqn:Einstein}
-\end{equation}
-
-Curabitur mi sem, pulvinar quis aliquam rutrum. (1) edf (2)
-, $\Omega=[-1,1]^3$, maecenas leo est, ornare at. $z=-1$ edf $z=1$ sed interdum felis dapibus sem. $x$ set $y$ ytruem. 
-Turpis $j$ amet accumsan enim $y$-lacina; 
-ref $k$-viverra nec porttitor $x$-lacina.
-
-Vestibulum ac diam a odio tempus congue. Vivamus id enim nisi:
-
-\begin{eqnarray}
-\cos\bar{\phi}_k Q_{j,k+1,t} + Q_{j,k+1,x}+\frac{\sin^2\bar{\phi}_k}{T\cos\bar{\phi}_k} Q_{j,k+1} &=&\nonumber\\ 
--\cos\phi_k Q_{j,k,t} + Q_{j,k,x}-\frac{\sin^2\phi_k}{T\cos\phi_k} Q_{j,k}\label{edgek}
-\end{eqnarray}
-and
-\begin{eqnarray}
-\cos\bar{\phi}_j Q_{j+1,k,t} + Q_{j+1,k,y}+\frac{\sin^2\bar{\phi}_j}{T\cos\bar{\phi}_j} Q_{j+1,k}&=&\nonumber \\
--\cos\phi_j Q_{j,k,t} + Q_{j,k,y}-\frac{\sin^2\phi_j}{T\cos\phi_j} Q_{j,k}.\label{edgej}
-\end{eqnarray}
-
-Nulla sed arcu arcu.
-Duis et ante gravida orci venenatis tincidunt.
-Fusce vitae lacinia metus.
-Pellentesque habitant morbi.
-$\mathbf{A}\underline{\xi}=\underline{\beta}$ Vim $\underline{\xi}$ enum nidi $3(P+2)^{2}$ lacina.
-Id feugain $\mathbf{A}$ nun quis; magno.
 
@@
 
-\section*{Source code}
+\subsection*{Source code}
 https://github.com/MugiwaraLuffy/ACN/

### 2013-12-17 poster/poster.tex
format poster

@@
 \usepackage{times} % Use the times font
-%\usepackage{palatino} % Uncomment to use the Palatino font
+\usepackage{palatino} % Uncomment to use the Palatino font
 
@@
 
-\color{DarkSlateGray} % DarkSlateGray color for the rest of the content
-
-\section*{Problem}
-
+\color{SaddleBrown}% DarkSlateGray color for the rest of the content
+\section*{PROBLEM}
+\color{DarkSlateGray}
 \begin{tabular} {l | l}
@@
     \item Optimality vs. I dont know what
-    \item Application design complexity (Architecture aware vs. agnostic)
+    \item Application design complexity (Architecture aware vs. agnostic)\\
 \end{itemize}
 
-\textbf{Our Goals}
+\textbf{Goals}
 \begin{itemize}
@@
     \item Enable smooth transition across management barriers
-    \item Handle changes locally
+    \item Handle changes locally \\
 \end{itemize}
@@
 
-\section*{Our solution}
+\color{SaddleBrown}
+\section*{PROPOSED SOLUTION}
+\color{DarkSlateGray}
 \begin{itemize}
@@
 %----------------------------------------------------------------------------------------
-
-\section*{Experimental set up}
-
+\color{SaddleBrown}
+\section*{EXPERIMENTAL SETUP}
+\color{DarkSlateGray}
 \begin{itemize}
@@
 
-\section*{Metrices}
+\color{SaddleBrown}
+\section*{METRICS}
+\color{DarkSlateGray}
 \begin{itemize}
@@
 
-\section*{Results}
+\color{SaddleBrown}
+\section*{RESULTS}
+\color{DarkSlateGray}
 Hierarchical architecture reduces communication overhead (Figure \ref{fig:msg}) by an order of magnitude while comprimising modestly on optimality: shortest path length (17.4\%) (Figure \ref{fig:path}) and minimum maximum link utilization (12.8\%) (Figure \ref{fig:util}).
@@
 
-%\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out
+\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out
 
-%\section*{Conclusions}
+%\section*{CONCLUSION}
 
@@
 %\section*{Forthcoming Research}
+
+\color{SaddleBrown}
+\subsection*{SOURCE CODE}
+\color{DarkSlateGray}
+https://github.com/MugiwaraLuffy/ACN/
 
@@
 
-\subsection*{Source code}
-https://github.com/MugiwaraLuffy/ACN/
-
 %----------------------------------------------------------------------------------------

### 2013-12-17 poster/poster.tex
added authors and logo
a1

@@
 
-\documentclass[a0,portrait]{a0poster}
+\documentclass[portrait,a1]{a0poster}
 
@@
 
-\usepackage{times} % Use the times font
+%x\usepackage{times} % Use the times font
 \usepackage{palatino} % Uncomment to use the Palatino font
@@
 \begin{minipage}[b]{0.75\linewidth}
-\veryHuge \color{NavyBlue} \textbf{Hierarchical Controller Architecture} \color{Black}\\ % Title
+\Huge \color{NavyBlue} \textbf{Hierarchical Controller Architecture} \color{Black}\\ % Title
 \Huge\color{NavyBlue}\text{in Software Defined Networks}\\[1cm] % Subtitle
-\huge \textbf{Author1 \& Author2}\\[0.5cm] % Author(s)
+\Large \textbf{Author1 \& Author2}\\[0.5cm] % Author(s)
 %\huge University and Department Name\\[0.4cm] % University/organization
@@
 %
-\begin{minipage}[b]{0.25\linewidth}
-\includegraphics[width=20cm]{logo.png}\\
-\end{minipage}
 
-\vspace{1cm} % A bit of extra whitespace between the header and poster content
+\vspace{-1cm} % A bit of extra whitespace between the header and poster content
 
@@
 \centering
-\includegraphics[scale=2]{hierarchy}
+\includegraphics[scale=1.9]{hierarchy}
 \caption{Hierarchical arrangement of controllers}
@@
 \centering
-\includegraphics[scale=1]{msg}
+\includegraphics[scale=01]{msg}
 \caption{Per flow communication overhead}
@@
   \begin{center}  
-    \subfigure[Path length]{\label{fig:path}\includegraphics[scale=0.75]{path.png}} 
-    \subfigure[Maximum link utilization]{\label{fig:util}\includegraphics[scale=0.75]{util.png}} 
+    \subfigure[Path length]{\label{fig:path}\includegraphics[scale=0.55]{path.png}} 
+    \subfigure[Maximum link utilization]{\label{fig:util}\includegraphics[scale=0.55]{util.png}} 
   \end{center}
@@
 
-\color{SaddleBrown}
-\subsection*{SOURCE CODE}
-\color{DarkSlateGray}
-https://github.com/MugiwaraLuffy/ACN/
-
  %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
-
+\color{SaddleBrown}
+\section*{RESOURCES}
+\color{DarkSlateGray}
+\section*{Source Code}
+https://github.com/MugiwaraLuffy/ACN/
 \nocite{*} % Print all references regardless of whether they were cited in the poster or not
-\bibliographystyle{plain} % Plain referencing style
+%\bibliographystyle{plain} % Plain referencing style
 \bibliography{sample} % Use the example bibliography file sample.bib

### 2013-12-17 poster/poster.tex
more formatting
small stuff
sorry - changed structure of text a lot

@@
 \usepackage[it]{subfigure}
+\usepackage{epstopdf}
+ \usepackage{tabularx}
 
@@
 
-\begin{minipage}[b]{0.75\linewidth}
-\Huge \color{NavyBlue} \textbf{Hierarchical Controller Architecture} \color{Black}\\ % Title
+\begin{minipage}[b]{0.85\linewidth}
+\Huge \color{NavyBlue} \textbf{Hierarchical Controller Architecture} \color{Black}\\[0.5cm] % Title
 \Huge\color{NavyBlue}\text{in Software Defined Networks}\\[1cm] % Subtitle
-\Large \textbf{Author1 \& Author2}\\[0.5cm] % Author(s)
+\Large \text{Gourav Khaneja, Sweta Seethamraju \& Praveen Kumar}\\[0.5cm] % Author(s)
+
 %\huge University and Department Name\\[0.4cm] % University/organization
@@
 %
+\begin{minipage}[b]{0.15\linewidth}
+\includegraphics[scale=0.8]{UIUC_logo.png}\\
+\end{minipage}
 
@@
 
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%\color{Navy} % Navy color for the abstract
-%
-%\begin{abstract}
-%
-%Sed fringilla tempus hendrerit.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Etiam ut elit sit amet metus lobortis consequat sit amet in libero.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Phasellus vel sem magna.
-Nunc at convallis urna.
-isus ante.
-Pellentesque condimentum dui.
-Etiam sagittis purus non tellus tempor volutpat.
-Donec et dui non massa tristique adipiscing.
-Quisque vestibulum eros eu.
-Phasellus imperdiet, tortor vitae congue bibendum, felis enim sagittis lorem, et volutpat ante orci sagittis mi.
-Morbi rutrum laoreet semper.
-Morbi accumsan enim nec tortor consectetur non commodo nisi sollicitudin.
-Proin sollicitudin.
-Pellentesque eget orci eros.
-Fusce ultricies, tellus et pellentesque fringilla, ante massa luctus libero, quis tristique purus urna nec nibh.
-%
-%\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	INTRODUCTION
-%----------------------------------------------------------------------------------------
-
-%\color{SaddleBrown} % SaddleBrown color for the introduction
-%
-%\section*{Introduction}
-%
-%Aliquam non lacus dolor, \textit{a aliquam quam} \cite{Smith:2012qr}.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Nulla in nibh mauris.
-Donec vel ligula nisi, a lacinia arcu.
-Sed mi dui, malesuada vel consectetur et, egestas porta nisi.
-Sed eleifend pharetra dolor, et dapibus est vulputate eu.
-\textbf{Integer faucibus elementum felis vitae fringilla.} In hac habitasse platea dictumst.
-Duis tristique rutrum nisl, nec vulputate elit porta ut.
-Donec sodales sollicitudin turpis sed convallis.
-Etiam mauris ligula, blandit adipiscing condimentum eu, dapibus pellentesque risus.
-%
-%\textit{Aliquam auctor}, metus id ultrices porta, risus enim cursus sapien, quis iaculis sapien tortor sed odio.
-Mauris ante orci, euismod vitae tincidunt eu, porta ut neque.
-Aenean sapien est, viverra vel lacinia nec, venenatis eu nulla.
-Maecenas ut nunc nibh, et tempus libero.
-Aenean vitae risus ante.
-Pellentesque condimentum dui.
-Etiam sagittis purus non tellus tempor volutpat.
-Donec et dui non massa tristique adipiscing.
-%
-%----------------------------------------------------------------------------------------
-%	OBJECTIVES
-%----------------------------------------------------------------------------------------
+\color{DarkSlateGray}
+Typical Control plane architecture in SDN involves multiple peer controllers, managing subset of network, but keeping a global view of the network.
 
 \color{SaddleBrown}% DarkSlateGray color for the rest of the content
-\section*{PROBLEM}
+\section*{PROBLEMS}
 \color{DarkSlateGray}
-\begin{tabular} {l | l}
-    \textbf{Single Controller Architecture} & \textbf{Distributed controller Architecture} \\
-    \hline
-    Responsiveness & Communication overhead \\
-       Scalability & Scalability of global network view \\
-       Reliability & Network state inconsistencies \\
-                   & Management barriers across networks
-\end{tabular}
 
-\textbf{Trade-offs}
 \begin{itemize}
-    \item Consistency vs. Responsiveness
-    \item Optimality vs. I dont know what
-    \item Application design complexity (Architecture aware vs. agnostic)\\
-\end{itemize}
-
-\textbf{Goals}
-\begin{itemize}
-    \item Make control applications agnostic to architecture
-    \item Reduce communication overhead
-    \item Minimize network state inconsistencies
-    \item Enable smooth transition across management barriers
-    \item Handle changes locally \\
-\end{itemize}
-
-\textbf{Price we pay}
-\begin{itemize}
-    \item Solution not always optimal
+    \item Replication of entire network state requires huge communication overhead: hinders scalability of the system
+    \item Network inconsistency degrades the performance of control application
+    \item Scalability of global network view itself
 \end{itemize}
@@
 \color{DarkSlateGray}
+
+\textbf{Switch Abstraction}
 \begin{itemize}
-    \item Hierarchical arrangement of controllers
-    \item One switch abstraction
-    \item Translate messages across hierarchical layers
+    \item A controller creates an abstract switch representing the underlying managed network.
+    \item A controller exposes "OpenFlow" API to manage the abstract switch by other (parent) controller.
+    \item OpenFlow messages are translated between parent controller and underlying network. \\
+\end{itemize}
+
+\textbf{Hierarchical arrangement}
+\begin{itemize}
+    \item Controllers are arranged in a tree structure over the network using switch abstraction.
+    \item Controllers do not keep a global view of network, but only the part which they control.
+    \item Local application and network events are shielded from upper layer controllers.
 \end{itemize}
@@
 
+\textbf{Pros} 
+Reduce replication overhead; Eliminate network inconsistency; Enable smooth transition across management barriers \\ \\
+\textbf{Price we pay}  \\
+Sub-optimality for inter-domain operations. How much is the deviation from optimality ?
+
 %------------------------------------------------
 
-%Gourav - I'm starting from here ownwards: 
+%Gourav - I'm starting from here onwards: 
 %----------------------------------------------------------------------------------------
@@
 \item Topology: 600 hosts connected through a network of 200 switches and 10 Gbps links. The cluster is divided into 10 equally-sized control domains.
-\item TE-cum-routing coutrol application: When a new flow arrives, it finds the path with minimum maximum utilization.
+\item TE-cum-routing control application: When a new flow arrives, it finds the path with minimum maximum utilization.
 If multiple such paths exist (which is frequently the case, due to a large network), shortest path (in term of number of hops) is chosen.
@@
 \item Maximum link utilization of the path assigned to flows.
-\item Communication overhead (number of messages exchanged between controllers) for seting up routes for flows.   
+\item Communication overhead (number of messages exchanged between controllers) for setting up routes for flows.   
 \end{itemize}
@@
 \color{DarkSlateGray}
-Hierarchical architecture reduces communication overhead (Figure \ref{fig:msg}) by an order of magnitude while comprimising modestly on optimality: shortest path length (17.4\%) (Figure \ref{fig:path}) and minimum maximum link utilization (12.8\%) (Figure \ref{fig:util}).
+Hierarchical architecture reduces communication overhead (Figure \ref{fig:msg}) by an order of magnitude while compromising modestly on optimality: shortest path length (17.4\%) (Figure \ref{fig:path}) and minimum maximum link utilization (12.8\%) (Figure \ref{fig:util}).
 
@@
 \color{DarkSlateGray}
-\section*{Source Code}
-https://github.com/MugiwaraLuffy/ACN/
+https://github.com/MugiwaraLuffy/ACN/ \\
+https://github.com/SWIMProjectUCB/SWIM
 \nocite{*} % Print all references regardless of whether they were cited in the poster or not

### 2013-12-17 poster/poster.tex
minor changes

@@
 \textbf{Pros}
-Reduce replication overhead; Eliminate network inconsistency; Enable smooth transition across management barriers \\ \\
+\begin{itemize}
+    \item Reduces replication overhead
+    \item Minimizes network inconsistency
+    \item Enables smooth transition across management barriers
+\end{itemize}
 \textbf{Price we pay}  \\
@@
 https://github.com/SWIMProjectUCB/SWIM
-\nocite{*} % Print all references regardless of whether they were cited in the poster or not
+%\nocite{*} % Print all references regardless of whether they were cited in the poster or not
 %\bibliographystyle{plain} % Plain referencing style
-\bibliography{sample} % Use the example bibliography file sample.bib
+%\bibliography{sample} % Use the example bibliography file sample.bib
 

### 2013-12-17 poster_new/conference_poster_4.tex
new design for poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (11/06/13)
%
% baposter Class Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[portrait,a1paper,fontscale=0.5]{baposter} % Adjust the font scale/size here

\usepackage{graphicx} % Required for including images
\graphicspath{{figures/}} % Directory in which figures are stored

\usepackage{amsmath} % For typesetting math
\usepackage{amssymb} % Adds new symbols to be used in math mode

\usepackage{booktabs} % Top and bottom rules for tables
\usepackage{enumitem} % Used to reduce itemize/enumerate spacing
\usepackage{palatino} % Use the Palatino font
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{wrapfig} % Allows wrapping text around tables and figures
%\usepackage{subfig}
\usepackage{float}
\usepackage[it]{subfigure}
\usepackage{multicol} % Required for multiple columns
\setlength{\columnsep}{1.5em} % Slightly increase the space between columns
\setlength{\columnseprule}{0mm} % No horizontal rule between columns

%\usepackage{tikz} % Required for flow chart
%\usetikzlibrary{shapes,arrows} % Tikz libraries required for the flow chart in the template

\newcommand{\compresslist}{ % Define a command to reduce spacing within itemize/enumerate environments, this is used right after \begin{itemize} or \begin{enumerate}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
}

\definecolor{lightblue}{rgb}{0.145,0.6666,1} % Defines the color used for content box headers
\definecolor{col1}{rgb}{0.9,0.9,0.9}
\definecolor{col2}{rgb}{0.9,0.9,0.9}
\begin{document}

\begin{poster}
{
headerborder=closed, % Adds a border around the header of content boxes
colspacing=1em, % Column spacing
%bgColorOne=white, % Background color for the gradient on the left side of the poster
bgColorOne=col1,
%bgColorTwo=white, % Background color for the gradient on the right side of the poster
bgColorTwo=col2,
borderColor=lightblue, % Border color
headerColorOne=black, % Background color for the header in the content boxes (left side)
headerColorTwo=lightblue, % Background color for the header in the content boxes (right side)
headerFontColor=white, % Text color for the header text in the content boxes
boxColorOne=white, % Background color of the content boxes
textborder=roundedleft, % Format of the border around content boxes, can be: none, bars, coils, triangles, rectangle, rounded, roundedsmall, roundedright or faded
eyecatcher=true, % Set to false for ignoring the left logo in the title and move the title left
headerheight=0.1\textheight, % Height of the header
headershape=roundedright, % Specify the rounded corner in the content box headers, can be: rectangle, small-rounded, roundedright, roundedleft or rounded
headerfont=\Large\bf\textsc, % Large, bold and sans serif font in the headers of content boxes
%textfont={\setlength{\parindent}{1.5em}}, % Uncomment for paragraph indentation
linewidth=2pt % Width of the border lines around content boxes
}
%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
%
{\includegraphics[height=4em]{logo.png}} % First university/lab logo on the left
{\bf\textsc{Hierarchical Controller Architecture in Software Defined Networks}\vspace{0.5em}} % Poster title
{\textsc{\{ Gourav Khaneja, Sweta Seethamraju \& Praveen Kumar \} \hspace{12pt}}} % Author names and institution
{\includegraphics[height=4em]{logo.png}} % Second university/lab logo on the right

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\headerbox{Problem}{name=problem,column=0,span=1,row=0}{

Typical Control plane architecture in SDN involves multiple peer controllers, managing subset of network, but keeping a global view of the network.

\begin{itemize}
    \item Replication of entire network state requires huge communication overhead: hinders scalability of the system
    \item Network inconsistency degrades the performance of control application
    \item Scalability of global network view itself
\end{itemize}

\vspace{0.3em} % When there are two boxes, some whitespace may need to be added if the one on the right has more content
}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\headerbox{Proposed Solution}{name=solution,column=1,span=2,row=0,span=2}{

\textbf{Switch Abstraction}
\begin{itemize}
    \item A controller creates an abstract switch representing the underlying managed network.
    \item A controller exposes "OpenFlow" API to manage the abstract switch by other (parent) controller.
    \item OpenFlow messages are translated between parent controller and underlying network. \\
\end{itemize}

\textbf{Hierarchical arrangement}
\begin{itemize}
    \item Controllers are arranged in a tree structure over the network using switch abstraction.
    \item Controllers do not keep a global view of network, but only the part which they control.
    \item Local application and network events are shielded from upper layer controllers.
\end{itemize}

\begin{center}
\includegraphics[scale=0.7]{hierarchy}
\captionof{figure}{Hierarchical arrangement of controllers}
\label{fig:hierarchy}
\end{center}

\textbf{Pros}
\begin{itemize}
    \item Reduces replication overhead
    \item Minimizes network inconsistency
    \item Enables smooth transition across management barriers
\end{itemize}
\textbf{Price we pay}  \\
Sub-optimality for inter-domain operations. How much is the deviation from optimality ?

}

\headerbox{EXPERIMENTAL SETUP}{name=exp,column=0,row=1,below=problem}{
\begin{itemize}
%\item We simulated a dummy control application on both hierarchical and flat controller architectures, with same network topology and flow arrival distribution.
\item Flow data (6000 MapReduce jobs) was generated by sampling historical Hadoop traces on a 600 machine cluster at Facebook (1 million jobs). A dummy MapReduce application choses directory, map and reduce nodes randomly for each job.
\item Topology: 600 hosts connected through a network of 200 switches and 10 Gbps links. The cluster is divided into 10 equally-sized control domains.
\item TE-cum-routing control application: When a new flow arrives, it finds the path with minimum maximum utilization. If multiple such paths exist (which is frequently the case, due to a large network), shortest path (in term of number of hops) is chosen.
\end{itemize}
}

\headerbox{METRICS}{name=metrics,column=1,span=2,row=2,below=solution}{
\begin{itemize}
\item Path length (number of hops) of the routes assigned to flows.
\item Maximum link utilization of the path assigned to flows.
\item Communication overhead (number of messages exchanged between controllers) for setting up routes for flows.   
\end{itemize}

\begin{center}
\includegraphics[scale=0.3]{msg}
\captionof{figure}{Per flow communication overhead}
\label{fig:msg}
\end{center}

\begin{minipage}{.5\textwidth}
\begin{center}
\includegraphics[scale=0.3]{path.png}
\captionof{figure}{Path length}
\label{fig:path}
\end{center}
\end{minipage}

\begin{minipage}{.5\textwidth}
\begin{center}
\includegraphics[scale=0.3]{util.png}
\captionof{figure}{Maximum link utilization}
\label{fig:util}
\end{center}
\end{minipage}

}

\headerbox{RESULTS}{name=results,column=0,row=3,below=exp}{

Hierarchical architecture reduces communication overhead (Figure \ref{fig:msg}) by an order of magnitude while compromising modestly on optimality: shortest path length (17.4\%) (Figure \ref{fig:path}) and minimum maximum link utilization (12.8\%) (Figure \ref{fig:util}).
}

\headerbox{RESOURCES}{name=resources,column=0,span=1,row=3,below=results}{

https://github.com/MugiwaraLuffy/ACN/ \\
https://github.com/SWIMProjectUCB/SWIM

}

\end{poster}

\end{document}

### 2013-12-17 poster_new/conference_poster_4.tex
new design for poster - format

@@
 textborder=roundedleft, % Format of the border around content boxes, can be: none, bars, coils, triangles, rectangle, rounded, roundedsmall, roundedright or faded
-eyecatcher=true, % Set to false for ignoring the left logo in the title and move the title left
+eyecatcher=false, % Set to false for ignoring the left logo in the title and move the title left
 headerheight=0.1\textheight, % Height of the header
@@
 {\textsc{\{ Gourav Khaneja, Sweta Seethamraju \& Praveen Kumar \} \hspace{12pt}}} % Author names and institution
-{\includegraphics[height=4em]{logo.png}} % Second university/lab logo on the right
+{\hspace{120pt} \includegraphics[height=8em]{UIUC_logo.png}} % Second university/lab logo on the right
 

### 2013-12-17 poster_new/conference_poster_4.tex
new design for poster - more formatting

@@
 
-\documentclass[portrait,a1paper,fontscale=0.5]{baposter} % Adjust the font scale/size here
+\documentclass[portrait,a1paper,fontscale=0.45]{baposter} % Adjust the font scale/size here
 
@@
 
-\headerbox{METRICS}{name=metrics,column=1,span=2,row=2,below=solution}{
+\headerbox{METRICS}{name=metrics,column=0,span=3,row=2,below=solution}{
 \begin{itemize}
@@
 \end{itemize}
-
+\begin{multicols}{3}
 \begin{center}
@@
 
-\begin{minipage}{.5\textwidth}
+\begin{minipage}{.3\textwidth}
 \begin{center}
@@
 \end{minipage}
+\end{multicols}
 
@@
 
-\headerbox{RESULTS}{name=results,column=0,row=3,below=exp}{
+\headerbox{RESULTS}{name=results,column=0,span=2,row=3,below=metrics}{
 
@@
 
-\headerbox{RESOURCES}{name=resources,column=0,span=1,row=3,below=results}{
+\headerbox{RESOURCES}{name=resources,column=2,span=1,row=3,below=metrics}{
 

### 2013-12-17 poster_new/conference_poster_4.tex
positioning

@@
 
-\headerbox{Problem}{name=problem,column=0,span=1,row=0}{
+\headerbox{Problem}{name=problem,column=0,row=0}{
 
@@
 
-\headerbox{Proposed Solution}{name=solution,column=1,span=2,row=0,span=2}{
+\headerbox{Proposed Solution}{name=solution,column=0,below=problem}{
 
@@
 
-\headerbox{EXPERIMENTAL SETUP}{name=exp,column=0,row=1,below=problem}{
+\headerbox{EXPERIMENTAL SETUP}{name=exp,column=1,row=0}{
 \begin{itemize}
@@
 
-\headerbox{METRICS}{name=metrics,column=0,span=3,row=2,below=solution}{
+\headerbox{METRICS}{name=metrics,column=1,below=exp}{
 \begin{itemize}
@@
 
-\headerbox{RESULTS}{name=results,column=0,span=2,row=3,below=metrics}{
+\headerbox{RESULTS}{name=results,column=1,below=metrics}{
 
@@
 
-\headerbox{RESOURCES}{name=resources,column=2,span=1,row=3,below=metrics}{
+\headerbox{RESOURCES}{name=resources,column=1,below=results}{
 

### 2013-12-17 poster_new/conference_poster_4.tex
new design for poster - more formatting 3

@@
 
-\documentclass[portrait,a1paper,fontscale=0.45]{baposter} % Adjust the font scale/size here
+\documentclass[portrait,a1paper,fontscale=0.4]{baposter} % Adjust the font scale/size here
 
@@
 {\textsc{\{ Gourav Khaneja, Sweta Seethamraju \& Praveen Kumar \} \hspace{12pt}}} % Author names and institution
-{\hspace{120pt} \includegraphics[height=8em]{UIUC_logo.png}} % Second university/lab logo on the right
+{\hspace{40pt} \includegraphics[height=8em]{UIUC_logo.png}} % Second university/lab logo on the right
 
@@
 \end{itemize}
-\begin{multicols}{3}
+
 \begin{center}
@@
 
-\begin{minipage}{.3\textwidth}
+\begin{multicols}{2}
+\begin{minipage}{.5\textwidth}
 \begin{center}
-\includegraphics[scale=0.3]{path.png}
+\includegraphics[scale=0.2]{path.png}
 \captionof{figure}{Path length}
@@
 
-\begin{minipage}{.3\textwidth}
+\begin{minipage}{.5\textwidth}
 \begin{center}
-\includegraphics[scale=0.3]{util.png}
+\includegraphics[scale=0.2]{util.png}
 \captionof{figure}{Maximum link utilization}

### 2013-12-17 poster_new/conference_poster_4.tex
results change

@@
 \headerbox{RESULTS}{name=results,column=1,below=metrics}{
-
-Hierarchical architecture reduces communication overhead (Figure \ref{fig:msg}) by an order of magnitude while compromising modestly on optimality: shortest path length (17.4\%) (Figure \ref{fig:path}) and minimum maximum link utilization (12.8\%) (Figure \ref{fig:util}).
+    \begin{itemize}
+        \item Reduces communication overhead by TODO (Figure \ref{fig:msg})
+        \item Increases shortest path length by 17.4\% (Figure \ref{fig:path})
+        \item Increases minimum maximum link utilization by 12.8\% (Figure \ref{fig:util}).
+    \end{itemize}
+Hierarchical architecture reduces communication overhead by an order of magnitude while compromising modestly on optimality
 }

### 2013-12-17 poster_new/conference_poster_4.tex
height adjustment (almost)

@@
 \item Flow data (6000 MapReduce jobs) was generated by sampling historical Hadoop traces on a 600 machine cluster at Facebook (1 million jobs).
-A dummy MapReduce application choses directory, map and reduce nodes randomly for each job.
+\item A dummy MapReduce application choses directory, map and reduce nodes randomly for each job.
 \item Topology: 600 hosts connected through a network of 200 switches and 10 Gbps links. The cluster is divided into 10 equally-sized control domains.
@@
 \begin{center}
-\includegraphics[scale=0.2]{path.png}
+\includegraphics[scale=0.22]{path.png}
 \captionof{figure}{Path length}
@@
 \begin{center}
-\includegraphics[scale=0.2]{util.png}
+\includegraphics[scale=0.22]{util.png}
 \captionof{figure}{Maximum link utilization}
@@
     \begin{itemize}
-        \item Reduces communication overhead by TODO (Figure \ref{fig:msg})
-        \item Increases shortest path length by 17.4\% (Figure \ref{fig:path})
+        \item Reduces communication overhead by 90.5\% (Figure \ref{fig:msg}).
+        \item Increases shortest path length by 17.4\% (Figure \ref{fig:path}).
         \item Increases minimum maximum link utilization by 12.8\% (Figure \ref{fig:util}).
     \end{itemize}
-Hierarchical architecture reduces communication overhead by an order of magnitude while compromising modestly on optimality
+Hierarchical architecture reduces communication overhead by an order of magnitude while compromising modestly on optimality.
 }
